---
title: "Merge LA refinery data sets"
author: "Jerry Wu"
date: "7/18/2023"
output:
  html_document:
    df_print: paged
abstract: ''
subtitle: ''
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(rgdal)
library(sf)
library(plotly)
library(units)
library(lwgeom)
library(circular)
library(raster)
select <- dplyr::select
```

```{r}
# load data sets
full_data <- tibble()
names <- c(ws = 'windSpeed', wd = 'windDirection', H2S = 'Value', QC.Code = 'qcCode',
           Monitor = 'siteName', MinDist = 'NEAR_DIST')
for (file in list.files('data/raw_csv')){
  new_data <- read.csv(paste0('data/raw_csv/', file)) %>% rename(any_of(names))
  print(sum(is.na(new_data$H2S)))
  full_data <- bind_rows(full_data, new_data)
}

remove(new_data)
gc()
```

# Compute date variables
```{r}
full_data <- full_data %>% 
  drop_na(c("wd", "ws")) %>% 
  mutate(DateTime = parse_date_time(DateTime, c('mdy HM', 'ymd HMS')),
         DateTime = as.POSIXct(DateTime, "%Y-%m-%d %H:%M:%OS", tz = 'UTC'))

full_data <- full_data %>% 
  mutate(day = as.POSIXct(format(DateTime, "%Y-%m-%d"), format="%Y-%m-%d", tz = 'UTC'))
 
# Remove observations before 2020
full_data <- full_data %>% filter(day >= '2020-01-01') %>% distinct()
 
full_data <- full_data %>% 
  mutate(yearmonth = format(day, "%Y-%m"),
         year = format(day, "%Y"),
         month = format(day, "%m"),
         weekday = relevel(factor(wday(full_data$day, label=TRUE), ordered = FALSE), ref = "Sun"))
```

```{r}
# Check for duplicate time
# full_data %>%
#    group_by(DateTime, Monitor) %>%
#    summarise(n = n()) %>%
#    filter(n > 1)
```
We observe multiple rows for the same date time and monitor. 
From a quick glance, this could be a conversion issue from Date.Time to DateTime, since the H2S measurements are different. 
In other occasions, multiple rows exist as complements, where one row contains data on the missing values of the other.
It also appears that some times we get different measurements for the same time when it's not daylight saving conversion days?

- example: January 31, G Street, wd and ws are recorded more than once.
```{r}
# full_data %>%
#     group_by(DateTime, Monitor) %>%
#     summarise(n = n()) %>%
#     filter(n > 1) %>%
#     mutate(day = as.POSIXct(format(DateTime, "%Y-%m-%d"), format="%Y-%m-%d", tz = 'UTC')) %>%
#     group_by(day, Monitor) %>%
#     summarise(n = n())
```


```{r}
# Deal with duplicate rows for same time

# For the 213 monitor it seems like the duplicate rows are complements to the other
#https://stackoverflow.com/questions/45515218/combine-rows-in-data-frame-containing-na-to-make-complete-row
# coalesce_by_column <- function(df) {
#   return(dplyr::coalesce(!!! as.list(df)))
# }
# 
# data_213 <- full_data %>%
#   filter(Monitor == '213th & Chico') %>%
#   group_by(DateTime) %>%
#   summarise_all(coalesce_by_column)
# 
# full_data <- full_data %>% 
#   filter(Monitor != '213th & Chico') %>%
#   rbind(data_213)
```


```{r}
glimpse(full_data)
```

```{r}
summary(full_data)
```
Some concerns for the data include:
For TorranceAir_H2SWind, there are actually three sites, but only one site contains ws/wd information
What's the difference between DateTime and Date/Time?
DateTime would be the time after converting
What's the difference between Date/Time and Date/Time.First? The latter is present in FirstMethodist
For missing H2S, it could be missing due to mismatch from merging, or flagged due to issues such as low detection (minimm detection limit: 0.4)

# Any additional cleaning
```{r}
# For Judson, it appeared that longitude had some significant figure issues
full_data <- full_data %>%
  mutate(longitude = case_when(Monitor == 'Judson' ~ -118.268128,
                               Monitor == 'West HS' ~ -118.3659363,
                               Monitor == 'Elm Ave' ~ -118.3396301,
                               Monitor == 'North HS' ~ -118.3312225,
                               Monitor == 'G Street' ~ -118.2818168,
                               Monitor == 'Guenser Park' ~ -118.313499,
                               Monitor == 'Harbor Park' ~ -118.286028,
                               .default = longitude),
         latitude = case_when(Monitor == 'West HS' ~ 33.84825516,
                               Monitor == 'Elm Ave' ~ 33.86428452,
                               Monitor == 'North HS' ~ 33.83779526,
                               Monitor == 'G Street' ~ 33.77785794,
                               Monitor == 'Guenser Park' ~ 33.869231,
                               Monitor == 'Harbor Park' ~ 33.786252,
                               .default = latitude))
```

```{r}
# For observations below MDL, set it to half of MDL
# As confirmed, MDL for all monitors is 0.4
# Check how many observations are below MDL
full_data %>%
  group_by() %>%
  summarise('Below MDL' = sum(full_data$H2S < 0.4, na.rm=T),
            'Proportion' = sum(full_data$H2S < 0.4, na.rm=T) / n())

full_data <- full_data %>%
  mutate(H2S = if_else(H2S < 0.4, 0.2, H2S))
```



# Compute Daily Max H2S
```{r}
# Compute daily max
# -Inf (all NA) is converted to NA for that day
H2S_dm <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))

full_data <- full_data %>%
  left_join(H2S_dm, by = c('day', 'Monitor'))
```

# Compute Monthly Avearge H2S
```{r}
H2S_ma <- H2S_dm %>%
  mutate(yearmonth = format(day, "%Y-%m")) %>%
  group_by(yearmonth, Monitor) %>%
  summarise(H2S_monthly_average = mean(H2S_daily_max, na.rm = TRUE))

full_data <- full_data %>%
  left_join(H2S_ma, by = c('yearmonth', 'Monitor'))

remove(H2S_ma, H2S_dm)
```

# Compute daily average H2S
```{r}
# Compute daily average
H2S_da <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_avg = mean(H2S, na.rm=TRUE))

full_data <- full_data %>%
  left_join(H2S_da, by = c('day', 'Monitor'))

remove(H2S_da)
```

# Compute wind sector
```{r}
full_data <- full_data %>%
  mutate(wd_sec = case_when(0 <= wd & wd < 90 ~ 'Q1',
                            90 <= wd & wd < 180 ~ 'Q2',
                            180 <= wd & wd < 270 ~ 'Q3',
                            270 <= wd & wd <= 360 ~ 'Q4'))
```

# Water treatment plants
```{r}
la_wrp <- read_sf('shapefiles/LA_WRP.shp', layer = 'LA_WRP')

CRS_UTM <- CRS("+proj=utm +zone=11 ellps=WGS84")

# Convert both water treatment plants and monitor coordinates to UTM
la_wrp <- st_transform(la_wrp, CRS_UTM)

monitors <- full_data %>%
  select(Monitor, latitude, longitude) %>%
  distinct() %>% 
  st_as_sf(coords = c('longitude', 'latitude')) %>% 
  st_set_crs(4326) %>% 
  st_transform(CRS_UTM)

# get nearest wrp for each monitor
monitor_wrp <- monitors %>% 
  mutate(wrp_name = la_wrp$CWP_NAM[st_nearest_feature(monitors, la_wrp)],
         dist_wrp = apply(st_distance(monitors, la_wrp), 1, min),
         capacity = la_wrp$cpcty_m[st_nearest_feature(monitors, la_wrp)])

# we also want to find the angle between the monitor and wrp
angles <- tibble(monitors$geometry, la_wrp$geometry[st_nearest_feature(monitors, la_wrp)]) %>%
  pivot_longer(cols = everything()) %>% 
  pull(value) %>% # extract coordinates only
  st_transform(4326) %>% # convert to lat/lon for function below
  st_geod_azimuth() %>%
  set_units('degrees') %>% # convert to degrees
  drop_units()

angles <- angles[c(T, F)] # keep only odd index, valid pairs
angles <- if_else(angles < 0, angles + 360, angles)

monitor_wrp$wrp_angle <- angles

full_data <- full_data %>%
  left_join(tibble(monitor_wrp) %>% select(-geometry), join_by('Monitor'))
```

# Dominguez Channel
```{r}
# Read in the shape file, it already has a CRS
st_read('shapefiles/DominguezChannel_Carson.shp')
d_channel <- read_sf('shapefiles/DominguezChannel_Carson.shp', layer = 'DominguezChannel_Carson')
d_channel <- st_transform(d_channel, CRS_UTM) # convert to UTM crs

monitor_d_channel <- monitors %>%
  mutate(dist_dc = apply(st_distance(monitors, d_channel), 1, min),
         dist_213 = c(drop_units(st_distance(monitors, 
                                             monitors[monitors$Monitor == '213th & Chico',]))))

full_data <- full_data %>%
  left_join(tibble(monitor_d_channel) %>% select(-geometry), join_by(Monitor))
```

```{r}
dist_plot <- d_channel %>% 
  ggplot() +
  geom_sf() +
  geom_sf(data = monitors) +
  geom_sf_label(data = monitors %>% mutate(dist_dc = apply(st_distance(monitors, d_channel), 1, min)), 
                aes(label = paste0(Monitor, ' ', round(dist_dc, 2)))) +
  theme_minimal()

dist_plot
```

```{r}
remove(dist_plot, la_wrp, monitor_wrp, monitor_d_channel, d_channel)
gc()
```


# Weather data from National Weather Service
```{r}
klax_data <- read_csv('data/KLAX_2020-2023.csv', show_col_types = FALSE) %>% 
  filter(!row_number() == 1)
ktoa_data <- read_csv('data/KTOA_2020-2023.csv', show_col_types = FALSE) %>% 
  filter(!row_number() == 1)
klgb_data <- read_csv('data/KLGB_2020-2023.csv', show_col_types = FALSE) %>% 
  filter(!row_number() == 1)
```

```{r}
# compute daily average temperature, humidity, and precipitation
klax_daily <- klax_data %>%
  group_by(Date) %>%
  summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T),
            avg_hum = mean(as.numeric(relative_humidity), na.rm=T),
            precip = precip_accum_24_hour[which(!is.na(precip_accum_24_hour))[1]]) %>%
  mutate(precip = coalesce(as.numeric(precip), 0),
         Date = parse_date_time(Date, c('m/d/y')),
         Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC'))

klgb_daily <- klgb_data %>%
  group_by(Date) %>%
  summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T),
            avg_hum = mean(as.numeric(relative_humidity), na.rm=T),
            precip = precip_accum_24_hour[which(!is.na(precip_accum_24_hour))[1]]) %>%
  mutate(precip = coalesce(as.numeric(precip), 0),
         Date = parse_date_time(Date, c('m/d/y')),
         Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC'))

ktoa_daily <- ktoa_data %>%
  group_by(Date) %>%
  summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T),
            avg_hum = mean(as.numeric(relative_humidity), na.rm=T)) %>%
  mutate(Date = parse_date_time(Date, c('m/d/y')),
         Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC'))

remove(klax_data, ktoa_data, klgb_data)

# notice that the Torrance station is missing 10 days of observations
# for that 10 days, use the long beach airport data as substitutes
ktoa_daily <- bind_rows(ktoa_daily, 
                        anti_join(klgb_daily,ktoa_daily, by = 'Date') %>% select(-precip))
```


```{r}
# map each monitor to closest monitor
weather_stations <- tibble(station = c('KLAX', 'KLGB', 'KTOA'),
                           latitude = c(33.93806, 33.81167, 33.803979),
                           longitude = c(-118.33194, -118.38889, -118.339432)) %>%
  st_as_sf(coords = c('longitude', 'latitude')) %>% 
  st_set_crs(4326)

weather_stations <- st_transform(weather_stations, CRS_UTM)

closest_temp_hum_station <- weather_stations$station[st_nearest_feature(monitors, weather_stations)]

closest_precip_station <- weather_stations$station[st_nearest_feature(monitors, weather_stations %>% filter(station != 'KTOA'))]

monitors_weather <- tibble(Monitor = monitors$Monitor, 
                           weather_station = closest_temp_hum_station,
                           closest_precip_station = closest_precip_station)

weather_full <- bind_rows(klax_daily %>% select(-precip) %>% 
                            mutate(weather_station = 'KLAX'),
                          klgb_daily %>% select(-precip) %>% 
                            mutate(weather_station = 'KLGB'),
                          ktoa_daily %>% 
                            mutate(weather_station = 'KTOA'))

precip_full <- bind_rows(klax_daily %>% select(Date, precip) %>% 
                            mutate(closest_precip_station = 'KLAX'),
                          klgb_daily %>% select(Date, precip) %>% 
                            mutate(closest_precip_station = 'KLGB'))

full_data <- full_data %>%
  left_join(monitors_weather, join_by('Monitor')) %>%
  left_join(weather_full, join_by('day' == 'Date', 'weather_station')) %>%
  left_join(precip_full, join_by('day' == 'Date', 'closest_precip_station'))

remove(klax_daily, ktoa_daily, klgb_daily, precip_full, weather_full, 
       monitors_weather, weather_stations)
gc()
```


# Well production data
```{r}
well_prod <- read_csv('data/LA Well August2023 Well Monthly Production.CSV', show_col_types = FALSE)
well_info <- read_csv('data/LA Well August2023 Well Headers.CSV', show_col_types = FALSE)
```

```{r}
well_prod <- well_prod %>%
  left_join(well_info, join_by(`API/UWI` == API14))

well_prod <- well_prod %>%
  select(`API/UWI`, `Monthly Production Date`, `Monthly Oil`, `Monthly Gas`, 
         `Monthly Water`, `Monthly BOE`, Days, `Operator Company Name.x`, 
         `Production Type.x`, `Well Status.x`, `Drill Type`, 
         `Surface Hole Latitude (WGS84)`, `Surface Hole Longitude (WGS84)`) %>%
  rename(lon = `Surface Hole Longitude (WGS84)`,
         lat = `Surface Hole Latitude (WGS84)`)

# project distinct well locations to UTM
well_location <- well_prod %>%
  select(`API/UWI`, lon, lat) %>%
  distinct() %>%
  st_as_sf(coords = c('lon', 'lat')) %>% 
  st_set_crs(4326) %>%
  st_transform(CRS_UTM) 

# distance to closest monitor for each well
monitor_well <- well_location %>%
  mutate(dist_to_closest_monitor = apply(st_distance(monitors, well_location), 2, min),
         `1km` = dist_to_closest_monitor<=1000,
         `2p5km` = dist_to_closest_monitor<=2500,
         `5km` = dist_to_closest_monitor<=5000)

well_prod <- well_prod %>%
  left_join(monitor_well %>% select(-geometry), join_by(`API/UWI`))
```

```{r}
# get monthly production data at different distance levels (monitor to well)
prod_data <- list(
    well_prod %>% 
      filter(`1km` == TRUE) %>% 
      group_by(`Monthly Production Date`) %>% 
      summarise(active_1km = n(),
                monthly_oil_1km = sum(`Monthly Oil`, na.rm = TRUE),
                monthly_gas_1km = sum(`Monthly Gas`, na.rm = TRUE),
                monthly_water_1km = sum(`Monthly Water`, na.rm = TRUE),
                monthly_boe_1km = sum(`Monthly BOE`, na.rm = TRUE))
    ,
    well_prod %>% 
      filter(`2p5km` == TRUE) %>% 
      group_by(`Monthly Production Date`) %>% 
      summarise(active_2p5km = n(),
            monthly_oil_2p5km = sum(`Monthly Oil`, na.rm = TRUE),
            monthly_gas_2p5km = sum(`Monthly Gas`, na.rm = TRUE),
            monthly_water_2p5km = sum(`Monthly Water`, na.rm = TRUE),
            monthly_boe_2p5km = sum(`Monthly BOE`, na.rm = TRUE))
    ,
    well_prod %>% 
      filter(`5km` == TRUE) %>% 
      group_by(`Monthly Production Date`) %>% 
      summarise(active_5km = n(),
            monthly_oil_5km = sum(`Monthly Oil`, na.rm = TRUE),
            monthly_gas_5km = sum(`Monthly Gas`, na.rm = TRUE),
            monthly_water_5km = sum(`Monthly Water`, na.rm = TRUE),
            monthly_boe_5km = sum(`Monthly BOE`, na.rm = TRUE))
  ) %>%
  reduce(inner_join, by = 'Monthly Production Date') %>%
  mutate(yearmonth = format(`Monthly Production Date`, '%Y-%m'))

full_data <- full_data %>%
  left_join(prod_data, join_by(yearmonth))

remove(monitor_well, well_info, prod_data, well_info, well_location, well_prod)
gc()
```

```{r}
# cleanup columns
full_data <- full_data  %>%
  select(-c('Date.Time', 'MDL', 'Unit', 'Averaging.Hour', 'Join_Count', 'siteId', 'unitName', 
            'qcName', 'opName', 'windSpeed_Unit', 'windDirection_Unit', 'opCode', 'Source',
            'Monthly Production Date', 'active_2p5km', 'monthly_oil_2p5km', 'monthly_gas_2p5km',
            'monthly_water_2p5km', 'monthly_boe_2p5km', 'active_5km', 'monthly_oil_5km', 
            'monthly_gas_5km', 'monthly_water_5km', 'monthly_boe_5km'))
```

# Downwind Indicator wrp Refinery
```{r}
# Create variable to indicate downwind wrt refinery
wind_diff <- abs(full_data$Converted_Angle - 180 - full_data$wd)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
full_data$downwind_ref <- as.integer(wind_diff <= 15)
```

# Downwind of wrp
```{r}
# Finally, check if wind direction is downwind w.r.t wrp
wind_diff <- abs(full_data$wrp_angle - full_data$wd)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

full_data$downwind_wrp <- as.integer(wind_diff <= 15)
```

# Compute UTM coordinates for monitors
```{r}
CRS_UTM <- CRS("+proj=utm +zone=11 ellps=WGS84")

monitors <- full_data %>%
  select(Monitor, latitude, longitude) %>%
  distinct() %>% 
  st_as_sf(coords = c('longitude', 'latitude')) %>% 
  st_set_crs(4326) %>% 
  st_transform(CRS_UTM)
monitors <- tibble(Monitor = monitors$Monitor,
                   mon_utm_x = st_coordinates(monitors)[,1],
                   mon_utm_y = st_coordinates(monitors)[,2])

full_data <- full_data %>%
  left_join(monitors, join_by(Monitor))
```

```{r}
# Compute daily average wd/ws
daily_full <- full_data %>%
  select(H2S_daily_max, H2S_daily_avg, H2S_monthly_average, ws, wd, latitude, 
         longitude, mon_utm_x, mon_utm_y, Monitor, MinDist, 
         Converted_Angle, Refinery, year, month, day, weekday, 
         active_1km, monthly_oil_1km, monthly_gas_1km, 
         dist_wrp, capacity, wrp_angle, dist_dc, dist_213,
         avg_temp, avg_hum, precip) %>%
  group_by(Monitor, day) %>%
  mutate(ws_avg = mean(ws, na.rm=TRUE),
         wd_avg = as.numeric(mean(circular(wd, units = 'degrees'), na.rm=TRUE))) %>%
  mutate(wd_avg = if_else(wd_avg < 0, wd_avg+360, wd_avg)) %>%
  ungroup() %>%
  rename(monitor_lat = latitude, monitor_lon = longitude)  %>%
  select(-c(wd, ws)) %>%
  distinct()
```

```{r}
# Get the downwind indicators for daily data
wind_diff <- abs(daily_full$Converted_Angle - 180 - daily_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
daily_full$daily_downwind_ref <- as.integer(wind_diff <= 15)

wind_diff <- abs(daily_full$wrp_angle - daily_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

daily_full$daily_downwind_wrp <- as.integer(wind_diff <= 15)
```


# Get monitor locations
```{r}
monitors_coord <- full_data %>%
  select(Monitor, latitude, longitude) %>%
  distinct()

monitors <- monitors_coord %>%
  select(Monitor)
  
coordinates(monitors_coord) <- ~ longitude + latitude
```

# Load in elevation data
```{r}
elevation <- raster('shapefiles/N33W119.hgt')

monitors <- cbind(monitors, extract(elevation, monitors_coord)) %>%
  as.data.frame() %>%
  rename(elevation = 2)
```

# Load in EVI data
```{r}
evi <- raster('shapefiles/MOD13Q1.006__250m_16_days_EVI_doy2022177_aid0001.tif')

monitors <- cbind(monitors, extract(evi, monitors_coord) * 0.0001) %>%
  as.data.frame() %>%
  rename(EVI = 3)
```

# Merge EVI and elevation to data
```{r}
full_data <- full_data %>%
  left_join(monitors, join_by(Monitor))

daily_full <- daily_full %>%
  left_join(monitors, join_by(Monitor))
```

# Merge Odor Complaint data
```{r}
odor <- read_csv('data/SCAQMD_odorcomplaints_2018_2022.csv')
la_county <- read_sf('shapefiles/Zip_Codes_(LA_County)/Zip_Codes_(LA_County).shp')
```

```{r}
# https://gis.stackexchange.com/questions/282750/identify-polygon-containing-point-with-r-sf-package
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(monitors_coord@coords), 
function(i) {st_point(as.numeric(monitors_coord@coords[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, CRS_UTM) # apply transformation to pnts sf
la_county_trans <- st_transform(la_county, CRS_UTM)      # apply transformation to polygons sf

# intersect and extract state name
monitors_coord@data$county <- apply(st_intersects(la_county_trans, pnts_trans, sparse = FALSE), 2, 
               function(col) { 
                  la_county_trans[which(col), ]$ZIPCODE
               })
```

```{r}
library(ggrepel)
# Plot results to double check
la_county_present <- la_county[la_county$ZIPCODE %in% unique(monitors_coord$county), ]

monitor_zip <- tibble(Monitor = monitors_coord@data$Monitor,
                      zip = monitors_coord@data$county,
                      longitude = monitors_coord@coords[,1],
                      latitude = monitors_coord@coords[,2])

monitor_zip_graph <- ggplot() +
  geom_sf(data = la_county_present, aes(fill = ZIPCODE)) +
  geom_point(data = monitor_zip, 
             aes(x = longitude, y = latitude))

monitor_zip_graph + 
  geom_label_repel(data = monitor_zip, aes(x = longitude, y = latitude, label = Monitor),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50') +
  theme_minimal()
```

```{r}
odor <- odor[odor$zip %in% unique(monitors_coord$county), ]

monitor_odor <- monitors_coord@data %>%
  left_join(odor %>% mutate(zip = as.character(zip)), join_by(county == zip)) %>%
  select(-`...1`)

full_data <- full_data %>%
  left_join(monitors_coord@data, join_by(Monitor))

daily_full <- daily_full %>%
  left_join(monitors_coord@data, join_by(Monitor))

full_data <- full_data %>%
  left_join(monitor_odor %>% select(Monitor, date, num_odor_complaints), join_by(Monitor, day == date))

full_data <- full_data %>%
  mutate(num_odor_complaints = coalesce(num_odor_complaints, 0))

daily_full <- daily_full %>%
  left_join(monitor_odor %>% select(Monitor, date, num_odor_complaints), join_by(Monitor, day == date))

daily_full <- daily_full %>%
  mutate(num_odor_complaints = coalesce(num_odor_complaints, 0))
```

```{r}
odor_graph <- odor %>%
  filter(zip %in% unique(monitors_coord$county)) %>%
  mutate(zip = as.character(zip)) %>%
   ggplot(aes(x=date, y=num_odor_complaints, group=zip, color=zip)) +
   geom_line() +
   labs(title = "Number of odor complaints over time", y = 'Number of odor complaints', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(odor_graph) %>% layout(dragmode = 'pan')
```

```{r}
odor_graph_b <- odor %>%
  filter(zip %in% unique(monitors_coord$county)) %>%
  mutate(zip = as.character(zip)) %>%
   ggplot(aes(x=date, y=num_odor_complaints, group=zip, color=zip)) +
   geom_line() +
   labs(title = "Number of odor complaints over time w/o outliers", y = 'Number of odor complaints', x = 'time') +
   scale_y_continuous(limits = c(0, 30)) +  
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(odor_graph_b) %>% layout(dragmode = 'pan')
```


```{r}
saveRDS(full_data, 'data/full_data.rds')
saveRDS(daily_full, 'data/daily_full.rds')
```

```{r}
remove(la_county, la_county_present, la_county_trans, monitor_odor, monitor_zip,
       monitor_zip_graph, monitors, monitor_coord, odor, odor_graph, odor_graph_b,
       pnts_sf, pnts_trans, elevation, evi, monitors_coord)
gc()
```


# Other issues
- Seems like for Torrance Refinery, the three monitors had different locations recorded at first, but stayed the same for many observations.
