---
title: "Data Visualizations"
author: "Jerry Wu"
date: "7/18/2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
abstract: ''
subtitle: ''
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(openairmaps)
library(htmlwidgets)
library(openair)
library(leaflet)
library(htmltools)
library(sf)
```

```{r}
full_data <- readRDS('data/full_data.rds')
daily_full <- readRDS('data/daily_full.rds')
```

# Monthly Production Data
```{r}
prod_by_month <- full_data %>% 
    select(yearmonth, active_1km, monthly_oil_1km, monthly_gas_1km, 
           monthly_water_1km, monthly_boe_1km) %>%
  distinct() %>%
  filter(yearmonth <= '2023-03')

coef <- median(1032*prod_by_month$monthly_gas_1km/10^9)/median(prod_by_month$monthly_oil_1km/10^6)

full_data %>%
  ggplot(aes(x = yearmonth, group = 1)) + 
  geom_line(aes(y = monthly_oil_1km/10^6, color = 'Oil'), size=1) + 
  geom_line(aes(y = 1032*monthly_gas_1km/10^9/coef, color = 'Natural Gas'), linewidth=1) +
  scale_y_continuous(
    name = "Monthly Oil in Million Barrels",
    sec.axis = sec_axis(~.*coef, name="Monthly Gas in Billion Cubic Feet")
  ) +
  theme(
    axis.title.y = element_text(color = 'tomato'),
    axis.title.y.right = element_text(color = 'skyblue')
  ) +
  scale_color_manual(values = c("skyblue", "tomato")) +
  scale_x_discrete(breaks = prod_by_month$yearmonth[seq(1, length(prod_by_month$yearmonth), by = 7)]) +
  xlab("Date") + labs(colour="Production Type") +
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

# Visualizing Daily Max H2S
```{r}
H2S_dm <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))
```

```{r}
H2S_ma <- H2S_dm %>%
  mutate(yearmonth = format(day, "%Y-%m")) %>%
  group_by(yearmonth, Monitor) %>%
  summarise(H2S_monthly_average = mean(H2S_daily_max, na.rm = TRUE))
```

```{r}
H2S_dm_graph <- H2S_dm %>%
 ggplot(aes(x=day, y=H2S_daily_max, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Max H2S concentration by monitor", y = 'Daily Max H2S', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_dm_graph_b <- H2S_dm %>%
 ggplot(aes(x=day, y=H2S_daily_max, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Max H2S concentration by monitor w/o outlier", y = 'Daily Max H2S', x = 'time') +
   scale_y_continuous(limits = c(0, 100)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph_b) %>% layout(dragmode = 'pan')
```





```{r}
H2S_ma_graph <- H2S_ma %>%
 ggplot(aes(x=yearmonth, y=H2S_monthly_average, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Monthly Average H2S concentration by monitor", y = 'Monthly Average H2S', x = 'time') +
   scale_x_discrete(breaks = unique(H2S_ma$yearmonth)[seq(1, length(unique(H2S_ma$yearmonth)), by = 6)]) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_ma_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_ma_graph_b <- H2S_ma %>%
 ggplot(aes(x=yearmonth, y=H2S_monthly_average, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Monthly Average H2S concentration by monitor w/o outlier", y = 'Monthly Average H2S', x = 'time') +
   scale_x_discrete(breaks = unique(H2S_ma$yearmonth)[seq(1, length(unique(H2S_ma$yearmonth)), by = 6)]) +
  scale_y_continuous(limits=c(0, 50)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_ma_graph_b) %>% layout(dragmode = 'pan')
```
The line with the largest deviation/peak corresponds to the 213th&Chico monitor, in 2021-10
Maybe we should start from January 2022 to start modelling to remove the extreme values, or simply removing the 213 monitor since that's set up specifically for the event.

# Visualizing Daily Average H2S
```{r}
# Compute daily average
H2S_da <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_avg = mean(H2S, na.rm=TRUE))
```

```{r}
H2S_da_graph <- H2S_da %>%
 ggplot(aes(x=day, y=H2S_daily_avg, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Average H2S concentration by monitor", y = 'Daily Average H2S', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_da_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_da_graph_b <- H2S_da %>%
 ggplot(aes(x=day, y=H2S_daily_avg, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Average H2S concentration by monitor w/o outlier", y = 'Daily Avearge H2S', x = 'time') +
   scale_y_continuous(limits = c(0, 100)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph_b) %>% layout(dragmode = 'pan')
```

```{r}
full_data %>%
  polarPlot(pollutant = "H2S", col = "turbo", 
            key.position = "bottom",
            key.header = "mean H2S", 
            key.footer = NULL, title = 'hi')
```

## Monitor/refinery Locations
```{r}
#https://atlas.eia.gov/datasets/6547eda91ef84cc386e23397cf834524_22/about
st_read('shapefiles/Petroleum_Refinery.shp')
refineries <- read_sf('shapefiles/Petroleum_Refinery.shp', layer = 'Petroleum_Refinery')
refineries <- refineries %>%
  select(Company, Site, Latitude, Longitude) %>%
  mutate(refinery = case_when(Site == 'Carson' ~ 'Marathon (Carson)',
                              Site == 'El Segundo' ~ 'Chevron El Segundo',
                              Site == 'Wilmington Asphalt Plant' ~ 'Valero Refinery',
                              Site == 'Wilmington Refinery' ~ 'Marathon (Wilmington)',
                              Site == 'Torrance' ~ 'Torrance Refinery',
                              Site == 'Wilmington' ~ 'Phillips 66 (Wilmington)'))
```


```{r}
monitor_loc <- daily_full %>%
  select(monitor_lon, monitor_lat, Monitor) %>%
  distinct()

ref_loc <- daily_full %>%
  select(Refinery) %>% distinct()

# also load in dominguez channel
st_read('shapefiles/DominguezChannel_Carson.shp')
d_channel <- read_sf('shapefiles/DominguezChannel_Carson.shp', layer = 'DominguezChannel_Carson')

basemap <- leaflet() %>%
              addProviderTiles(providers$MtbMap) %>%
              addProviderTiles(providers$Stamen.TerrainBackground,
                               options = providerTileOptions(opacity = 0.9)) %>%
              addProviderTiles(providers$Stamen.TonerLines,
                options = providerTileOptions(opacity = 0.25)) %>%
              addProviderTiles(providers$Stamen.TonerLabels) %>%
              addCircleMarkers(data = monitor_loc, 
                         lng=~monitor_lon, lat=~monitor_lat, 
                         fillOpacity = 0.85, radius = 4,
                         fillColor = , weight = 1, stroke = T, col = '#BA55D3',
                         label = ~Monitor) %>%
              addCircleMarkers(data = refineries, 
                         lng=~Longitude, lat=~Latitude, 
                         fillOpacity = 0.95, radius = 6,
                         fillColor = , weight = 1, stroke = T, col = '#8B0000',
                         label = ~refinery)


basemap
saveWidget(basemap, file="basemap.html")
```


## Since Feb 2022
```{r}
# Create a polar map
# TBD: include markers for refineries
polarmap1 <- polarMap(
  daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01',
           !(Monitor %in% c('West HS', 'Elm Ave', 'North HS'))) %>%
    rename(wd = wd_avg,
           ws = ws_avg),
  pollutant = "H2S_daily_avg",
  latitude = "monitor_lat",
  longitude = "monitor_lon",
  popup = "Monitor",
  provider = "Stamen.Toner",
  d.icon = 150,
  d.fig = 2.5,
  alpha = 0.95,
  key = TRUE
)

polarmap2 <- polarMap(
  daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01',
           Monitor %in% c('West HS', 'Elm Ave', 'North HS')) %>%
    rename(wd = wd_avg,
           ws = ws_avg),
  pollutant = "H2S_daily_avg",
  latitude = "monitor_lat",
  longitude = "monitor_lon",
  popup = "Monitor",
  provider = "Stamen.Toner",
  d.icon = 150,
  d.fig = 2.5,
  alpha = 0.95,
  key = TRUE,
  statistic = 'nwr'
)

polarmap1

polarmap2
```

```{r}
sincefeb2022_mon <- daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01') %>% pull(Monitor) %>% unique()

sincefeb2022_data <- daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01') %>%
    rename(wd = wd_avg,
           ws = ws_avg)

for (mon in sincefeb2022_mon) {
  polarPlot(sincefeb2022_data %>% filter(Monitor == mon),
    pollutant = "H2S_daily_avg",
    type = 'Monitor',
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 30)
}
```


```{r}
polarPlot(daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01',
           Monitor == 'West HS') %>%
    rename(wd = wd_avg,
           ws = ws_avg),
    pollutant = "H2S_daily_avg",
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 10)

polarPlot(daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01',
           Monitor == 'North HS') %>%
    rename(wd = wd_avg,
           ws = ws_avg),
    pollutant = "H2S_daily_avg",
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 10)

polarPlot(daily_full[complete.cases(daily_full),] %>%
    filter(day >= '2022-02-01',
           Monitor == 'Elm Ave') %>%
    rename(wd = wd_avg,
           ws = ws_avg),
    pollutant = "H2S_daily_avg",
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 10)
```

## Disaster Period
```{r}
dis_mon <- daily_full[complete.cases(daily_full),] %>%
    filter(year == '2021', month %in% c('10', '11', '12')) %>% pull(Monitor) %>% unique()

dis_data <- daily_full[complete.cases(daily_full),] %>%
    filter(year == '2021', month %in% c('10', '11', '12')) %>%
    rename(wd = wd_avg,
           ws = ws_avg)

for (mon in dis_mon) {
  polarPlot(dis_data %>% filter(Monitor == mon),
    pollutant = "H2S_daily_avg",
    type = 'Monitor',
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 30)
}
```

```{r}
# save these plots
getplot <- function(mon) {
  mon_plot <- polarPlot(dis_data %>% filter(Monitor == mon),
    pollutant = "H2S_daily_avg",
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 30,
    key = FALSE)
  
  return(mon_plot)
}

# for loop doesn't work with png() and bg='transparent' for some reason...
mon <- dis_mon[1]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[2]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[3]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[4]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[5]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()


mon <- dis_mon[6]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[7]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[8]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[9]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[10]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[11]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[12]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- dis_mon[13]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/disaster/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()


parOrig <- par()
par(parOrig)
plot (1:3)
```

## Since 2020

### Full
```{r}
polarmap3 <- polarMap(
  daily_full[complete.cases(daily_full),] %>%
    rename(wd = wd_avg,
           ws = ws_avg),
  pollutant = "H2S_daily_avg",
  latitude = "monitor_lat",
  longitude = "monitor_lon",
  popup = "Monitor",
  provider = "Stamen.Toner",
  d.icon = 150,
  d.fig = 2.5,
  alpha = 0.95,
  key = TRUE
)

polarmap3
```

```{r}
polarPlot(daily_full[complete.cases(daily_full),] %>%
    rename(wd = wd_avg,
           ws = ws_avg) %>% 
      filter(Monitor == '213th & Chico'),
    pollutant = "H2S_daily_avg",
    type = 'Monitor',
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 30)
```
```{r}
full_mon <- daily_full[complete.cases(daily_full),] %>% 
  pull(Monitor) %>% unique()

full_data <- daily_full[complete.cases(daily_full),] %>%
    rename(wd = wd_avg,
           ws = ws_avg)

for (mon in full_mon) {
  polarPlot(full_data %>% filter(Monitor == mon),
    pollutant = "H2S_daily_avg",
    type = 'Monitor',
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    k = 30)
}
```


### Normal period
```{r}
polarmap4 <- polarMap(
  daily_full[complete.cases(daily_full),] %>%
    filter(!(year == '2021' & month %in% c('10', '11', '12'))) %>%
    rename(wd = wd_avg,
           ws = ws_avg),
  pollutant = "H2S_daily_avg",
  latitude = "monitor_lat",
  longitude = "monitor_lon",
  popup = "Monitor",
  provider = "Stamen.Toner",
  d.icon = 150,
  d.fig = 2.5,
  alpha = 0.95,
  key = TRUE
)

polarmap4
```

```{r}
norm_mon <- daily_full[complete.cases(daily_full),] %>% 
  filter(!(year == '2021' & month %in% c('10', '11', '12'))) %>%
  pull(Monitor) %>% unique()

norm_data <- daily_full[complete.cases(daily_full),] %>%
  filter(!(year == '2021' & month %in% c('10', '11', '12'))) %>%
  rename(wd = wd_avg,
           ws = ws_avg)

# 213 needs lower k due to lack of data
polar_norm_213 <- polarPlot(norm_data %>% filter(Monitor == '213th & Chico'),
  pollutant = "H2S_daily_avg",
  latitude = "monitor_lat",
  longitude = "monitor_lon",
  k = 10,
  key = FALSE)

png("polarplots/normal/213th_&_Chico.png", width=20, height=15, units="cm", res=800, bg="transparent")
polar_norm_213
dev.off()

plot(polar_norm_213, bg = 'transparent')
dev.copy(png,'myplot.png',  bg = 'transparent')

# for the rest, use k=30
getplot <- function(mon) {
  mon_plot <- polarPlot(norm_data %>% filter(Monitor == mon),
    pollutant = "H2S_daily_avg",
    latitude = "monitor_lat",
    longitude = "monitor_lon",
    #limits = c(0, 1.5),
    k = 30,
    key = FALSE)
  
  return(mon_plot)
}

# for loop doesn't work with png() and bg='transparent' for some reason...
mon <- norm_mon[2]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[3]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[4]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[5]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()


mon <- norm_mon[6]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[7]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[8]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[9]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[10]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[11]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[12]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()

mon <- norm_mon[13]
mon_plot <- getplot(mon)
mon <- str_replace_all(str_replace_all(mon, '[()]', ''), ' ', '_')
png(paste0("polarplots/normal/", mon, ".png"), width=20, height=15, units="cm", res=800, bg="transparent")
mon_plot
dev.off()


parOrig <- par()
par(parOrig)
plot (1:3)
```

# Visualize Odor Complaint data
```{r}
odor <- daily_full %>%
  select(day, county, num_odor_complaints) %>%
  distinct()

odor_graph <- odor %>%
   ggplot(aes(x=day, y=num_odor_complaints, group=county, color=county)) +
   geom_line() +
   labs(title = "Number of odor complaints over time", y = 'Number of odor complaints', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(odor_graph) %>% layout(dragmode = 'pan')
```

```{r}
odor_graph_b <- odor %>%
   ggplot(aes(x=day, y=num_odor_complaints, group=county, color=county)) +
   geom_line() +
   labs(title = "Number of odor complaints over time", y = 'Number of odor complaints', x = 'time') +
  scale_y_continuous(limits = c(0, 30)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(odor_graph_b) %>% layout(dragmode = 'pan')
```