---
title: "Merge LA refinery data ses"
author: "Jerry Wu"
date: "5/1/2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
abstract: ''
subtitle: ''
---

```{r setup, include=FALSE}
library(tidyverse)
library(mgcv)
library(plotly)
library(tidyquant)
library(openair)
library(caret)
library(xgboost)
library(fastDummies)
library(openairmaps)
library(htmlwidgets)
```

```{r}
full_data <- readRDS('data/full_data.rds')
```


# Compute Daily Max H2S
```{r}
H2S_dm <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))
```

```{r}
H2S_ma <- H2S_dm %>%
  mutate(yearmonth = format(day, "%Y-%m")) %>%
  group_by(yearmonth, Monitor) %>%
  summarise(H2S_monthly_average = mean(H2S_daily_max, na.rm = TRUE))
```

```{r}
H2S_dm_graph <- H2S_dm %>%
 ggplot(aes(x=day, y=H2S_daily_max, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Max H2S concentration by monitor", y = 'Daily Max H2S', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_dm_graph_b <- H2S_dm %>%
 ggplot(aes(x=day, y=H2S_daily_max, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Max H2S concentration by monitor w/o outlier", y = 'Daily Max H2S', x = 'time') +
   scale_y_continuous(limits = c(0, 100)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph_b) %>% layout(dragmode = 'pan')
```





```{r}
H2S_ma_graph <- H2S_ma %>%
 ggplot(aes(x=yearmonth, y=H2S_monthly_average, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Monthly Average H2S concentration by monitor", y = 'Monthly Average H2S', x = 'time') +
   scale_x_discrete(breaks = unique(H2S_ma$yearmonth)[seq(1, length(unique(H2S_ma$yearmonth)), by = 6)]) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_ma_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_ma_graph_b <- H2S_ma %>%
 ggplot(aes(x=yearmonth, y=H2S_monthly_average, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Monthly Average H2S concentration by monitor w/o outlier", y = 'Monthly Average H2S', x = 'time') +
   scale_x_discrete(breaks = unique(H2S_ma$yearmonth)[seq(1, length(unique(H2S_ma$yearmonth)), by = 6)]) +
  scale_y_continuous(limits=c(0, 50)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_ma_graph_b) %>% layout(dragmode = 'pan')
```
The line with the largest deviation/peak corresponds to the 213th&Chico monitor, in 2021-10
Maybe we should start from January 2022 to start modelling to remove the extreme values, or simply removing the 213 monitor since that's set up specifically for the event.

# Daily average
```{r}
# Compute daily average
H2S_da <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_avg = mean(H2S, na.rm=TRUE))
```

```{r}
H2S_da_graph <- H2S_da %>%
 ggplot(aes(x=day, y=H2S_daily_avg, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Average H2S concentration by monitor", y = 'Daily Average H2S', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_da_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_da_graph_b <- H2S_da %>%
 ggplot(aes(x=day, y=H2S_daily_avg, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Average H2S concentration by monitor w/o outlier", y = 'Daily Avearge H2S', x = 'time') +
   scale_y_continuous(limits = c(0, 100)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph_b) %>% layout(dragmode = 'pan')
```

```{r}

full_data %>%
  polarPlot(pollutant = "H2S", col = "turbo", 
            key.position = "bottom",
            key.header = "mean H2S", 
            key.footer = NULL, title = 'hi')
```

```{r}
# Create a polar map
# TBD: include markers for refineries
polarplot <- polarMap(
  full_data %>% 
    filter(!(is.na(latitude) | is.na(H2S) | is.na(wd) | is.na(ws))) %>%
    rename(date = DateTime),
  pollutant = "H2S",
  latitude = "latitude",
  longitude = "longitude",
  popup = "Monitor",
  provider = "Esri.WorldImagery",
  d.icon = 150,
  d.fig = 2.5,
  alpha = 0.75
)
polarplot

#saveWidget(polarplot, file="polarplot.html")
```


```{r}
# Compute daily average wd/ws
daily_full <- full_data %>%
              select(c(H2S, ws, wd, latitude, longitude, Monitor, MinDist, 
                       Converted_Angle, Refinery, year, month, day, weekday)) %>%
  group_by(Monitor, day, latitude, longitude, Refinery, Converted_Angle, year, 
           month, MinDist, weekday) %>%
  summarise(H2S_daily_max = max(H2S, na.rm=TRUE),
            H2S_daily_avg = mean(H2S, na.rm=TRUE),
            ws_avg = mean(ws, na.rm=TRUE),
            wd_avg = mean(wd, na.rm=TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))
```

# Explore some GAM models
### Five minute H2S
```{r}
h2s_model_a <- gam(H2S~s(as.numeric(month),bs='cc')+year+wd_sec+ws+I(1/MinDist^2)+Refinery, data = full_data)
summary(h2s_model_a)
plot(h2s_model_a)
```

```{r}
h2s_model_b <- gam(H2S~s(as.numeric(month),bs='cc')+wd_sec+ws+I(1/MinDist^2)+
                   Refinery+s(latitude, longitude, bs='tp', k = 5), 
                   data = full_data %>% filter(day >= '2022-02-01'))
summary(h2s_model_b)
plot(h2s_model_b)
```

```{r}
# Include converted angle and weekday
h2s_model_c <- gam(H2S ~ s(as.numeric(month),bs='cc') + year + wd_sec + ws + 
                     I(1/MinDist^2) + Converted_Angle + Refinery + 
                     s(latitude, longitude, bs='tp', k = 10) + 
                     as.factor(weekday), 
                   data = full_data %>% filter(day >= '2022-02-01'))
summary(h2s_model_c)
plot(h2s_model_c)
```

### Daily max H2S
```{r}
# With month, year, wind sector/speed, dist to refinery, refinery
h2s_dm_model_a <- gam(H2S_daily_max~s(as.numeric(month),bs='cc')+year+wd_avg+ws_avg+
                        I(1/MinDist^2)+Refinery, 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_a)
plot(h2s_dm_model_a)
```

```{r}
# Also with location of monitor
h2s_dm_model_b <- gam(H2S_daily_max~s(as.numeric(month),bs='cc')+wd_avg+ws_avg+
                        I(1/MinDist^2)+Refinery+s(latitude, longitude, bs='tp', k=10), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_b)
plot(h2s_dm_model_b)
```

```{r}
# Also include angle to refinery
h2s_dm_model_c <- gam(H2S_daily_max~s(as.numeric(month),bs='cc')+wd_avg+ws_avg+
                        I(1/MinDist^2)+Refinery+Converted_Angle+
                        s(latitude, longitude, bs='tp', k=10), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_c)
plot(h2s_dm_model_c)
```
### Daily Average
```{r}
# Look at daily average
h2s_da_model_a <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc')+wd_avg+ws_avg+
                        I(1/MinDist^2)+Refinery+Converted_Angle+
                        s(latitude, longitude, bs='tp', k=10), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_a)
plot(h2s_da_model_a)
```

```{r}
# Include year too, since we have 2023 now
h2s_da_model_b <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc')+wd_avg+ws_avg+year+
                        I(1/MinDist^2)+Refinery+Converted_Angle+
                        s(latitude, longitude, bs='tp', k=10), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_b)
plot(h2s_da_model_b)
```

# Monthly 
```{r}
h2s_ma_model_a <- gam(H2S_monthly_average~s(as.numeric(month),bs='cc')+year+wd_sec+ws+I(1/MinDist^2)+Refinery, data = full_data)
summary(h2s_ma_model_a)
plot(h2s_ma_model_a)
```


```{r}
h2s_ma_model_b <- gam(H2S_monthly_average~s(as.numeric(month),bs='cc')+year+wd_sec+ws+I(1/MinDist^2)+Refinery+s(latitude, longitude, bs='tp', k=10), data = full_data)
summary(h2s_ma_model_b)
plot(h2s_ma_model_b)
```

# Try XGBoost on Daily Max

```{r echo = T, results = 'hide'}
daily_full <- daily_full %>%
  mutate(Refinery = str_replace_all(str_replace_all(Refinery, '[()]', ''), ' ', '_'),
         Monitor = str_replace_all(Refinery, ' ', '_'),
         weekday = as.character(weekday))

## Try for a continuous month
# tune_grid <- expand.grid(nrounds = c(100, 200),
#                          max_depth = c(2, 3, 4),
#                          eta = c(0.1, 0.3),
#                          gamma = c(0.01, 0.001),
#                          colsample_bytree = c(0.5, 1),
#                          min_child_weight = 0,
#                          subsample = c(0.5, 0.75, 1))
# # Run algorithms using 10-fold cross validation
# control <- trainControl(method="cv", number=10,verboseIter=TRUE, search='grid')

fit.xgb_da <- readRDS('rfiles/fit.xgb_da.rds')
# fit.xgb_da <- train(H2S_daily_avg~., 
#                  method = 'xgbTree', 
#                  data = fastDummies::dummy_cols(daily_full[complete.cases(daily_full),] %>% 
#                    ungroup() %>% 
#                    filter(day >= '2022-02-01') %>% 
#                    select(-c(Refinery, latitude, longitude, day, H2S_daily_max)) %>%
#                    mutate(MinDist = 1/(MinDist^2)), 
#                    remove_selected_columns = TRUE)
#                  ,
#                  trControl=control,
#                  tuneGrid = tune_grid,
#                  tuneLength = 10, importance=TRUE, verbosity = 0, verbose=FALSE)
# saveRDS(fit.xgb_da, 'rfiles/fit.xgb_da.rds')
```

```{r}
getTrainPerf(fit.xgb_da)
```

```{r}
fit.xgb_da$finalModel
```


```{r}
imp<-varImp(fit.xgb_da,scale=FALSE)
plot(imp, top=10)
```

## SHAP for XGBoost above (Caret)
```{r}
matrix <- fastDummies::dummy_cols(daily_full[complete.cases(daily_full),] %>% 
                   ungroup() %>% 
                   filter(day >= '2022-02-01') %>% 
                   select(-c(Refinery, latitude, longitude, day, H2S_daily_max, H2S_daily_avg)) %>%
                   mutate(MinDist = 1/(MinDist^2)), 
                   remove_selected_columns = TRUE)
matrix <- as.matrix(sapply(matrix, as.numeric))  
xgb.plot.shap(data = matrix, 
              model = fit.xgb_da$finalModel, top_n = 12, n_col = 3)
```

```{r}
xgb.ggplot.shap.summary(data = matrix, 
              model = fit.xgb_da$finalModel, top_n = 10)
```