---
title: "Data Analysis and modelling"
author: "Jerry Wu"
date: "5/1/2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
abstract: ''
subtitle: ''
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(mgcv)
library(mgcViz)
library(plotly)
library(tidyquant)
library(openair)
library(caret)
library(xgboost)
library(fastDummies)
library(openairmaps)
library(htmlwidgets)
library(circular)
library(raster)
library(geosphere)
library(terra)
library(rgl)
select <- dplyr::select
```

```{r}
full_data <- readRDS('data/full_data.rds')
```

# Monthly Production Data
```{r}
prod_by_month <- full_data %>% 
    select(yearmonth, active_1km, monthly_oil_1km, monthly_gas_1km, 
           monthly_water_1km, monthly_boe_1km, active_2p5km, monthly_oil_2p5km, 
           monthly_gas_2p5km, monthly_water_2p5km, monthly_boe_2p5km, active_5km, 
           monthly_oil_5km, monthly_gas_5km, monthly_water_5km, monthly_boe_5km) %>%
  distinct() %>%
  filter(yearmonth <= '2023-03')

coef <- median(1032*prod_by_month$monthly_oil_1km/10^9)/median(prod_by_month$monthly_gas_1km/10^6)

full_data %>%
  ggplot(aes(x = yearmonth, group = 1)) + 
  geom_line(aes(y = monthly_oil_1km/10^6, color = 'Oil'), size=1) + 
  geom_line(aes(y = 1032*monthly_gas_1km/10^9/coef, color = 'Natural Gas'), size=1) +
  scale_y_continuous(
    name = "Monthly Oil in Million Barrels",
    sec.axis = sec_axis(~.*coef, name="Monthly Gas in Billion Cubic Feet")
  ) +
  theme(
    axis.title.y = element_text(color = 'tomato'),
    axis.title.y.right = element_text(color = 'skyblue')
  ) +
  scale_color_manual(values = c("skyblue", "tomato")) +
  scale_x_discrete(breaks = prod_by_month$yearmonth[seq(1, length(prod_by_month$yearmonth), by = 7)]) +
  xlab("Date") + labs(colour="Production Type") +
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

# Visualizing Daily Max H2S
```{r}
H2S_dm <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))
```

```{r}
H2S_ma <- H2S_dm %>%
  mutate(yearmonth = format(day, "%Y-%m")) %>%
  group_by(yearmonth, Monitor) %>%
  summarise(H2S_monthly_average = mean(H2S_daily_max, na.rm = TRUE))
```

```{r}
H2S_dm_graph <- H2S_dm %>%
 ggplot(aes(x=day, y=H2S_daily_max, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Max H2S concentration by monitor", y = 'Daily Max H2S', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_dm_graph_b <- H2S_dm %>%
 ggplot(aes(x=day, y=H2S_daily_max, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Max H2S concentration by monitor w/o outlier", y = 'Daily Max H2S', x = 'time') +
   scale_y_continuous(limits = c(0, 100)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph_b) %>% layout(dragmode = 'pan')
```





```{r}
H2S_ma_graph <- H2S_ma %>%
 ggplot(aes(x=yearmonth, y=H2S_monthly_average, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Monthly Average H2S concentration by monitor", y = 'Monthly Average H2S', x = 'time') +
   scale_x_discrete(breaks = unique(H2S_ma$yearmonth)[seq(1, length(unique(H2S_ma$yearmonth)), by = 6)]) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_ma_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_ma_graph_b <- H2S_ma %>%
 ggplot(aes(x=yearmonth, y=H2S_monthly_average, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Monthly Average H2S concentration by monitor w/o outlier", y = 'Monthly Average H2S', x = 'time') +
   scale_x_discrete(breaks = unique(H2S_ma$yearmonth)[seq(1, length(unique(H2S_ma$yearmonth)), by = 6)]) +
  scale_y_continuous(limits=c(0, 50)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_ma_graph_b) %>% layout(dragmode = 'pan')
```
The line with the largest deviation/peak corresponds to the 213th&Chico monitor, in 2021-10
Maybe we should start from January 2022 to start modelling to remove the extreme values, or simply removing the 213 monitor since that's set up specifically for the event.

# Visualizing Daily Average H2S
```{r}
# Compute daily average
H2S_da <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_avg = mean(H2S, na.rm=TRUE))
```

```{r}
H2S_da_graph <- H2S_da %>%
 ggplot(aes(x=day, y=H2S_daily_avg, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Average H2S concentration by monitor", y = 'Daily Average H2S', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_da_graph) %>% layout(dragmode = 'pan')
```

```{r}
H2S_da_graph_b <- H2S_da %>%
 ggplot(aes(x=day, y=H2S_daily_avg, group=Monitor, color=Monitor)) +
   geom_line() +
   labs(title = "Daily Average H2S concentration by monitor w/o outlier", y = 'Daily Avearge H2S', x = 'time') +
   scale_y_continuous(limits = c(0, 100)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(H2S_dm_graph_b) %>% layout(dragmode = 'pan')
```

```{r}
full_data %>%
  polarPlot(pollutant = "H2S", col = "turbo", 
            key.position = "bottom",
            key.header = "mean H2S", 
            key.footer = NULL, title = 'hi')
```

```{r}
# Create a polar map
# TBD: include markers for refineries
polarplot <- polarMap(
  full_data %>% 
    filter(!(is.na(latitude.x) | is.na(H2S) | is.na(wd) | is.na(ws))) %>%
    rename(date = DateTime),
  pollutant = "H2S",
  latitude = "latitude.x",
  longitude = "longitude.x",
  popup = "Monitor",
  provider = "Esri.WorldImagery",
  d.icon = 150,
  d.fig = 2.5,
  alpha = 0.75
)
polarplot

#saveWidget(polarplot, file="polarplot.html")
```

```{r}
# Create variable to indicate downwind wrt refinery
wind_diff <- abs(full_data$Converted_Angle - 180 - full_data$wd)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
full_data$downwind_ref <- as.integer(wind_diff <= 15)
```

# Find angle between monitor and nearest water treatment plant
```{r}
# Function to calculate the angle between two lat/long points
calculate_angle <- function(lat1, lon1, lat2, lon2) {

  # Calculate the bearing between the two points
  angle <- bearing(cbind(lon1, lat1), cbind(lon2, lat2))
  
  # Return the angle
  return(angle)
}

wrp <- full_data %>% 
  select(name, wrp_utm_x, wrp_utm_y) %>%
  distinct()

wrp_coord <- cbind(wrp, terra::project(as.matrix(cbind(wrp$wrp_utm_x, wrp$wrp_utm_y)), 
                            "+proj=utm +zone=11 ellps=WGS84", 
                            "+proj=longlat +datum=WGS84")) %>%
  rename(longitude.wrp = '1',
         latitude.wrp = '2')

full_data <- full_data %>%
  left_join(wrp_coord %>% select(name, latitude.wrp, longitude.wrp), join_by(name), keep=F)

mon_wrp <- full_data %>% 
              select(Monitor, name, latitude.x, longitude.x, latitude.wrp, longitude.wrp) %>%
              distinct()

mon_wrp$wrp_angle <- calculate_angle(mon_wrp$latitude.x, mon_wrp$longitude.x, 
                                     mon_wrp$latitude.wrp, mon_wrp$longitude.wrp)

mon_wrp$wrp_angle <- if_else(mon_wrp$wrp_angle < 0, mon_wrp$wrp_angle + 360, mon_wrp$wrp_angle)

full_data <- full_data %>%
  left_join(mon_wrp %>% select(Monitor, name, wrp_angle), join_by(Monitor, name), keep=F)

# Finally, check if wind direction is downwind w.r.t wrp
wind_diff <- abs(full_data$wrp_angle - full_data$wd)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

full_data$downwind_wrp <- as.integer(wind_diff <= 15)
```


```{r}
# Compute daily average wd/ws
daily_full <- full_data %>%
  select(H2S, ws, wd, latitude.x, longitude.x, mon_utm_x, mon_utm_y, Monitor, MinDist, 
         Converted_Angle, Refinery, year, month, day, weekday, 
         monthly_oil_1km, monthly_gas_1km, monthly_oil_2p5km, 
         monthly_gas_2p5km, monthly_oil_5km, monthly_gas_5km,
         dist_wrp, capacity, active_1km, active_2p5km, active_5km,
         wrp_angle) %>%
  group_by(Monitor, day) %>%
  mutate(H2S_daily_max = max(H2S, na.rm=TRUE),
            H2S_daily_avg = mean(H2S, na.rm=TRUE),
            ws_avg = mean(ws, na.rm=TRUE),
            wd_avg = as.numeric(mean(circular(wd, units = 'degrees'), na.rm=TRUE))) %>%
  mutate(wd_avg = if_else(wd_avg < 0, wd_avg+360, wd_avg)) %>%
  ungroup() %>%
  rename(monitor_lat = latitude.x, monitor_lon = longitude.x) %>%
  mutate(H2S_daily_max = ifelse(H2S_daily_max == -Inf, NA, H2S_daily_max)) %>%
  select(-c(H2S, wd, ws)) %>%
  distinct()
```

```{r}
# Get the downwind indicators for daily data
wind_diff <- abs(daily_full$Converted_Angle - 180 - daily_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
daily_full$daily_downwind_ref <- wind_diff

wind_diff <- abs(daily_full$wrp_angle - daily_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

daily_full$daily_downwind_wrp <- as.integer(wind_diff <= 15)
```


# Get monitor locations
```{r}
monitors_coord <- full_data %>%
  select(Monitor, latitude.x, longitude.x) %>%
  distinct()

monitors <- monitors_coord %>%
  select(Monitor)
  
coordinates(monitors_coord) <- ~ longitude.x + latitude.x
```

# Load in elevation data
```{r}
elevation <- raster('shapefiles/N33W119.hgt')

monitors <- cbind(monitors, extract(elevation, monitors_coord)) %>%
  as.data.frame() %>%
  rename(elevation = 2)
```

# Load in EVI data
```{r}
evi <- raster('shapefiles/MOD13Q1.006__250m_16_days_EVI_doy2022177_aid0001.tif')

monitors <- cbind(monitors, extract(evi, monitors_coord) * 0.0001) %>%
  as.data.frame() %>%
  rename(EVI = 3)
```

# Merge EVI and elevation to data
```{r}
full_data <- full_data %>%
  left_join(monitors, join_by(Monitor))

daily_full <- daily_full %>%
  left_join(monitors, join_by(Monitor))
```

# Explore some GAM models
### Five minute H2S
```{r}
h2s_model_a <- gam(H2S~s(as.numeric(month),bs='cc')+year+wd_sec+ws+I(1/MinDist^2), 
                   data = full_data %>% filter(day >= '2022-02-01'))
summary(h2s_model_a)
plot(h2s_model_a)
```

```{r}
h2s_model_b <- gam(H2S~s(as.numeric(month),bs='cc')+wd_sec+ws+I(1/MinDist^2)+
                   s(latitude.x, longitude.x, bs='tp', k = 8), 
                   data = full_data %>% filter(day >= '2022-02-01'))
summary(h2s_model_b)
plot(h2s_model_b)
```

```{r}
# Include converted angle and weekday
h2s_model_c <- gam(H2S ~ s(as.numeric(month),bs='cc') + year + wd_sec + ws + 
                     I(1/MinDist^2) + Converted_Angle + 
                     s(latitude.x, longitude.x, bs='tp', k = 10) + 
                     as.factor(weekday), 
                   data = full_data %>% filter(day >= '2022-02-01'))
summary(h2s_model_c)
plot(h2s_model_c)
```

```{r}
# Include converted angle and weekday
h2s_model_d <- gam(H2S ~ 
                     s(as.numeric(month),bs='cc') + year + as.factor(weekday) +
                     wd_sec + ws + downwind_ref + downwind_wrp +
                     I(1/MinDist^2) + dist_wrp + capacity +
                     Converted_Angle + 
                     s(latitude.x, longitude.x, bs='tp', k = 10) + 
                     monthly_oil_1km + monthly_gas_1km + active_1km
                     , 
                   data = full_data %>% filter(day >= '2022-02-01'))
summary(h2s_model_d)
plot(h2s_model_d)
```

### Daily max H2S
```{r}
# With month, year, wind sector/speed, dist to refinery, refinery
h2s_dm_model_a <- gam(H2S_daily_max~s(as.numeric(month),bs='cc')+year+wd_avg+ws_avg+
                        I(1/MinDist^2), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_a)
plot(h2s_dm_model_a)
```

```{r}
# Also with location of monitor
h2s_dm_model_b <- gam(H2S_daily_max~s(as.numeric(month),bs='cc')+wd_avg+ws_avg+
                        I(1/MinDist^2)+Refinery+s(monitor_lat, monitor_lon, bs='tp', k=10), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_b)
plot(h2s_dm_model_b)
```

```{r}
# Also include angle to refinery
h2s_dm_model_c <- gam(H2S_daily_max~s(as.numeric(month),bs='cc')+wd_avg+ws_avg+
                        I(1/MinDist^2)+Converted_Angle+
                        s(monitor_lat, monitor_lon, bs='tp', k=10), 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_c)
plot(h2s_dm_model_c)
```

```{r}
h2s_dm_model_d <- gam(H2S_daily_max ~ 
                         s(as.numeric(month),bs='cc') + year + as.factor(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                         I(1/MinDist^2) + dist_wrp + capacity +
                         Converted_Angle + 
                         s(monitor_lat, monitor_lon, bs='tp', k = 10) + 
                         monthly_oil_1km + monthly_gas_1km + active_1km
                      , 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_dm_model_d)
plot(h2s_dm_model_d)
```

```{r}
# Try to include as many variables as possible
h2s_dm_model_e <- gam(H2S_daily_max ~ s(as.numeric(month),bs='cc') + year + as.character(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                          I(1/dist_wrp^2) +
                         s(I(mon_utm_x/10^3), I(mon_utm_y/10^3), bs='tp', k = 10) + 
                        te(I(mon_utm_x/10^3), I(mon_utm_y/10^3),as.numeric(day),k=c(10,10),d=c(2,1),bs=c('tp','cc')) +
                         monthly_oil_1km + monthly_gas_1km + active_1km + 
                        daily_downwind_wrp + elevation + EVI, 
                      data = daily_full %>% filter(day >= '2022-02-01'))

summary(h2s_dm_model_e)
plot(h2s_dm_model_e)

e <- getViz(h2s_dm_model_e)
plot(sm(e, 2)) + l_fitRaster() + l_fitContour() + l_points()

plotRGL(sm(e, 3), fix = c(`as.numeric(day)` = 0), residuals = F)

# try a plot using wireframe
plot3d <- matrix(fitted(h2s_dm_model_e), ncol = 20)
wireframe(plot3d, drape=TRUE, colorkey=TRUE)
```


### Daily Average
```{r}
# Look at daily average
h2s_da_model_a <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc') + year + as.factor(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                         I(1/MinDist^2) + I(1/dist_wrp^2) + capacity +
                         Converted_Angle + 
                         s(monitor_lon, monitor_lat, bs='tp', k = 10) + 
                         monthly_oil_1km + monthly_gas_1km + active_1km, 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_a)
plot(h2s_da_model_a)
a <- getViz(h2s_da_model_a)
plot(sm(a, 2)) + l_fitRaster() + l_fitContour() + l_points()
```

```{r}
# Look at daily average
h2s_da_model_b <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc') + year + as.factor(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                         I(1/MinDist^2) + I(1/dist_wrp^2) + capacity +
                         Converted_Angle + 
                         s(I(mon_utm_x/10^3), I(mon_utm_y/10^3), bs='tp', k = 10) + 
                         monthly_oil_1km + monthly_gas_1km + active_1km, 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_b)
plot(h2s_da_model_b)
b <- getViz(h2s_da_model_b)
plot(sm(b, 2)) + l_fitRaster() + l_fitContour() + l_points()
```

```{r}
h2s_da_model_c <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc') + year + as.factor(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                          I(1/dist_wrp^2) +
                          
                         s(I(mon_utm_x/10^3), I(mon_utm_y/10^3), bs='tp', k = 10) + 
                         monthly_oil_1km + monthly_gas_1km + active_1km, 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_c)
plot(h2s_da_model_c)
c <- getViz(h2s_da_model_c)
plot(sm(c, 2)) + l_fitRaster() + l_fitContour() + l_points()
```

```{r}
h2s_da_model_d <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc') + year + as.character(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                          I(1/dist_wrp^2) +
                          
                         s(I(mon_utm_x/10^3), I(mon_utm_y/10^3), bs='tp', k = 10) + 
                         monthly_oil_1km + monthly_gas_1km + active_1km + 
                        daily_downwind_wrp + elevation + EVI, 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_d)
plot(h2s_da_model_d)
d <- getViz(h2s_da_model_d)
plot(sm(d, 2)) + l_fitRaster() + l_fitContour() + l_points()
```

```{r}
h2s_da_model_e <- gam(H2S_daily_avg~s(as.numeric(month),bs='cc') + year + as.character(weekday) +
                         wd_avg + ws_avg + daily_downwind_ref + 
                          I(1/dist_wrp^2) +
                         s(I(mon_utm_x/10^3), I(mon_utm_y/10^3), bs='tp', k = 10) + 
                        te(I(mon_utm_x/10^3), I(mon_utm_y/10^3),as.numeric(day),k=c(10,10),d=c(2,1),bs=c('tp','cc')) +
                         monthly_oil_1km + monthly_gas_1km + active_1km + 
                        daily_downwind_wrp + elevation + EVI, 
                      data = daily_full %>% filter(day >= '2022-02-01'))
summary(h2s_da_model_e)
plot(h2s_da_model_e)

d <- getViz(h2s_da_model_e)
plot(sm(d, 2)) + l_fitRaster() + l_fitContour() + l_points()

plotRGL(sm(d, 3), fix = c(`as.numeric(day)` = 0), residuals = F)

# try a plot using wireframe
plot3d <- matrix(fitted(h2s_da_model_e), ncol = 20)
wireframe(plot3d, drape=TRUE, colorkey=TRUE)
```

# Monthly 
```{r}
h2s_ma_model_a <- gam(H2S_monthly_average~s(as.numeric(month),bs='cc')+year+wd_sec+ws+I(1/MinDist^2)+Refinery, data = full_data)
summary(h2s_ma_model_a)
plot(h2s_ma_model_a)
```


```{r}
h2s_ma_model_b <- gam(H2S_monthly_average~s(as.numeric(month),bs='cc')+year+wd_sec+ws+I(1/MinDist^2)+Refinery+s(latitude.x, longitude.x, bs='tp', k=10), data = full_data)
summary(h2s_ma_model_b)
plot(h2s_ma_model_b)
```



# Try XGBoost on Daily Average

```{r echo = T, results = 'hide'}
# Cross validation by leaving one monitor out 
daily_full <- daily_full %>%
  mutate(Refinery = str_replace_all(str_replace_all(Refinery, '[()]', ''), ' ', '_'),
         Monitor = str_replace_all(Monitor, ' ', '_'),
         weekday = as.character(weekday),
         daily_downwind_ref = as.integer(daily_downwind_ref),
         daily_downwind_wrp = as.integer(daily_downwind_wrp))

train <- daily_full %>%
  filter(Monitor != 'Harbor_Park')

train <- train[complete.cases(train),] %>%
  filter(day >= '2022-02-01')

train <- fastDummies::dummy_cols(train %>%
                   select(-c(Refinery, Monitor, day, active_2p5km, active_5km, monthly_oil_2p5km,
                             monthly_gas_2p5km, monthly_oil_5km, monthly_gas_5km, H2S_daily_max,
                             mon_utm_x, mon_utm_y)) %>%
                   mutate(MinDist = 1/(MinDist^2)),
                   remove_selected_columns = TRUE)

test <- daily_full %>%
  filter(Monitor == 'Harbor_Park')

test <- fastDummies::dummy_cols(test[complete.cases(test),] %>%
                   ungroup() %>%
                   filter(day >= '2022-02-01') %>%
                   select(-c(Refinery, Monitor, day, active_2p5km, active_5km, monthly_oil_2p5km,
                             monthly_gas_2p5km, monthly_oil_5km, monthly_gas_5km, H2S_daily_max,
                             mon_utm_x, mon_utm_y)) %>%
                   mutate(MinDist = 1/(MinDist^2)),
                   remove_selected_columns = TRUE)

# Try for a continuous month
tune_grid <- expand.grid(nrounds = c(100, 200, 500),
                         max_depth = c(3, 4, 5),
                         eta = c(0.1, 0.3),
                         gamma = c(0.01, 0.001),
                         colsample_bytree = c(0.5, 1),
                         min_child_weight = 0,
                         subsample = c(0.5, 0.75, 1))

# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", 
                        number=10,
                        verboseIter=TRUE, 
                        search='grid')

fit.xgb_da <- readRDS('rfiles/fit.xgb_da.rds')
# fit.xgb_da <- train(H2S_daily_avg~.,
#                  method = 'xgbTree',
#                  data = train,
#                  trControl=control,
#                  tuneGrid = tune_grid,
#                  tuneLength = 10, importance=TRUE, verbosity = 0, verbose=FALSE)
# saveRDS(fit.xgb_da, 'rfiles/fit.xgb_da.rds')
```

```{r}
getTrainPerf(fit.xgb_da)
```

```{r}
# Get testing performance on Harbor Park monitor
predictions <- predict(fit.xgb_da$finalModel, 
                       newdata = test %>% select(-H2S_daily_avg) %>% as.matrix())
R2(test[complete.cases(test),] %>% pull(H2S_daily_avg), predictions)
```

```{r}
fit.xgb_da$finalModel
```


```{r}
imp<-varImp(fit.xgb_da,scale=FALSE)
plot(imp, top=10)
```

## SHAP for XGBoost above (Caret)
```{r}
matrix <- test %>% select(-H2S_daily_avg)
matrix <- as.matrix(sapply(matrix, as.numeric))  
xgb.plot.shap(data = matrix, 
              model = fit.xgb_da$finalModel, top_n = 12, n_col = 3)
```

```{r}
xgb.ggplot.shap.summary(data = matrix, 
              model = fit.xgb_da$finalModel, top_n = 10)
```

# CV by leaving each monitor out once
```{r}
post2022feb <- daily_full %>% 
  filter(day >= '2022-02-01')

monitor_names <- unique(post2022feb$Monitor)

# for (i in 1:length(monitor_names)) {
#   train <- post2022feb %>%
#     filter(Monitor != monitor_names[i])
#   
#   train <- train[complete.cases(train),]
# 
#   train <- fastDummies::dummy_cols(train %>%
#                      select(-c(Refinery, Monitor, day, active_2p5km, active_5km, monthly_oil_2p5km,
#                                monthly_gas_2p5km, monthly_oil_5km, monthly_gas_5km, H2S_daily_max,
#                                mon_utm_x, mon_utm_y)) %>%
#                      mutate(MinDist = 1/(MinDist^2)),
#                      remove_selected_columns = TRUE)
# 
#   fit.xgb_da <- train(H2S_daily_avg~.,
#                  method = 'xgbTree',
#                  data = train,
#                  trControl=control,
#                  tuneGrid = tune_grid,
#                  tuneLength = 10, importance=TRUE, verbosity = 0, verbose=FALSE)
#   
#   saveRDS(fit.xgb_da, paste0('rfiles/fit.xgb_da_', monitor_names[i], '.rds'))
# }
```

```{r}
model_stats <- tibble(Monitor = character(), train_r2 = numeric(), test_r2 = numeric())

add_cols <- function(df, cols) {
  add <- cols[!cols %in% names(df)]
  if(length(add) !=0 ) df[add] <- NA
  return(df)
}

for (i in setdiff(1:length(monitor_names), c(9))) {

  fit.xgb_da <- readRDS(paste0('rfiles/fit.xgb_da_', monitor_names[i], '.rds'))
  
  test <- post2022feb %>%
    filter(Monitor == monitor_names[i])

  test <- fastDummies::dummy_cols(test[complete.cases(test),] %>%
                     ungroup() %>%
                     select(-c(Refinery, Monitor, day, active_2p5km, active_5km, monthly_oil_2p5km,
                           monthly_gas_2p5km, monthly_oil_5km, monthly_gas_5km, H2S_daily_max,
                           mon_utm_x, mon_utm_y)) %>%
                       add_cols(c('year_2022', 'year_2023', 'month_01', 'month_02',
                                  'month_03', 'month_04', 'month_05', 'month_06', 
                                  'month_07', 'month_08', 'month_09', 'month_10',
                                  'month_11', 'month_12')) %>%
                     mutate(MinDist = 1/(MinDist^2),
                            year_2022 = ifelse(is.na(year_2022), 0, year_2022),
                            year_2023 = ifelse(is.na(year_2023), 0, year_2023),
                            month_01 = ifelse(is.na(month_01), 0, month_01),
                            month_02 = ifelse(is.na(month_02), 0, month_02),
                            month_03 = ifelse(is.na(month_03), 0, month_03),
                            month_04 = ifelse(is.na(month_04), 0, month_04),
                            month_05 = ifelse(is.na(month_05), 0, month_05),
                            month_06 = ifelse(is.na(month_06), 0, month_06),
                            month_07 = ifelse(is.na(month_07), 0, month_07),
                            month_08 = ifelse(is.na(month_08), 0, month_08),
                            month_09 = ifelse(is.na(month_09), 0, month_09),
                            month_10 = ifelse(is.na(month_10), 0, month_10),
                            month_11 = ifelse(is.na(month_11), 0, month_11),
                            month_12 = ifelse(is.na(month_12), 0, month_12)),
                     remove_selected_columns = TRUE)
  
  train_r2 <- getTrainPerf(fit.xgb_da)$TrainRsquared
  
  predictions <- predict(fit.xgb_da$finalModel, 
                         newdata = test %>% select(-H2S_daily_avg) %>% as.matrix())
  
  test_r2 <- R2(test %>% pull(H2S_daily_avg), predictions)
  
  model_stats <- rbind(model_stats, tibble(Monitor = monitor_names[i], train_r2, test_r2))
}

model_stats %>% rbind(tibble(Monitor = 'Total Average', train_r2 = mean(model_stats$train_r2, na.rm = T),
                             test_r2 = mean(model_stats$test_r2, na.rm = T)))
```
