---
title: "Merge LA refinery data sets"
author: "Jerry Wu"
date: "5/1/2023"
output:
  html_document:
    df_print: paged
abstract: ''
subtitle: ''
---

```{r setup, include=FALSE}
library(tidyverse)
library(rgdal)
library(sf)
```

```{r}
# load data sets
full_data <- tibble()
names <- c(ws = 'windSpeed', wd = 'windDirection', H2S = 'Value', QC.Code = 'qcCode',
           Monitor = 'siteName', MinDist = 'NEAR_DIST')
for (file in list.files('data/raw_csv')){
  new_data <- read.csv(paste0('data/raw_csv/', file)) %>% rename(any_of(names))
  print(sum(is.na(new_data$H2S)))
  full_data <- bind_rows(full_data, new_data)
}
```

```{r}
full_data <- full_data %>% mutate(DateTime = parse_date_time(full_data$DateTime, c('mdy HM', 'ymd HMS')),
                                  DateTime = as.POSIXct(DateTime, "%Y-%m-%d %H:%M:%OS", tz = 'UTC'))

full_data <- full_data %>% 
  mutate(day = as.POSIXct(format(DateTime, "%Y-%m-%d"), format="%Y-%m-%d", tz = 'UTC'))
 
# Remove observations before 2020
full_data <- full_data %>% filter(day >= '2020-01-01') %>% distinct()
 
full_data <- full_data %>% 
  mutate(yearmonth = format(day, "%Y-%m"),
         year = format(day, "%Y"),
         month = format(day, "%m"),
         weekday = wday(full_data$day, label=TRUE))
```

```{r}
# Check for duplicate time
full_data %>%
   group_by(DateTime, Monitor) %>%
   summarise(n = n()) %>%
   filter(n > 1)
```
We observe multiple rows for the same date time and monitor. 
From a quick glance, this could be a conversion issue from Date.Time to DateTime, since the H2S measurements are different. 
In other occasions, multiple rows exist as complements, where one row contains data on the missing values of the other.
It also appears that some times we get different measurements for the same time when it's not daylight saving conversion days?

- example: January 31, G Street, wd and ws are recorded more than once.
```{r}
full_data %>%
    group_by(DateTime, Monitor) %>%
    summarise(n = n()) %>%
    filter(n > 1) %>%
    mutate(day = as.POSIXct(format(DateTime, "%Y-%m-%d"), format="%Y-%m-%d", tz = 'UTC')) %>%
    group_by(day, Monitor) %>%
    summarise(n = n())
```


```{r}
# Deal with duplicate rows for same time

# For the 213 monitor it seems like the duplicate rows are complements to the other
#https://stackoverflow.com/questions/45515218/combine-rows-in-data-frame-containing-na-to-make-complete-row
# coalesce_by_column <- function(df) {
#   return(dplyr::coalesce(!!! as.list(df)))
# }
# 
# data_213 <- full_data %>%
#   filter(Monitor == '213th & Chico') %>%
#   group_by(DateTime) %>%
#   summarise_all(coalesce_by_column)
# 
# full_data <- full_data %>% 
#   filter(Monitor != '213th & Chico') %>%
#   rbind(data_213)
```


```{r}
glimpse(full_data)
```

```{r}
summary(full_data)
```
Some concerns for the data include:
For TorranceAir_H2SWind, there are actually three sites, but only one site contains ws/wd information
What's the difference between DateTime and Date/Time?
DateTime would be the time after converting
What's the difference between Date/Time and Date/Time.First? The latter is present in FirstMethodist
For missing H2S, it could be missing due to mismatch from merging, or flagged due to issues such as low detection (minimm detection limit: 0.4)

# Any additional cleaning
```{r}
# For Judson, it appeared that longitude had some significant figure issues
full_data <- full_data %>%
  mutate(longitude = case_when(Monitor == 'Judson' ~ -118.268128,
                               Monitor == 'West HS' ~ -118.3659363,
                               Monitor == 'Elm Ave' ~ -118.3396301,
                               Monitor == 'North HS' ~ -118.3312225,
                               Monitor == 'G Street' ~ -118.2818168,
                               Monitor == 'Guenser Park' ~ -118.313499,
                               Monitor == 'Harbor Park' ~ -118.286028,
                               .default = longitude),
         latitude = case_when(Monitor == 'West HS' ~ 33.84825516,
                               Monitor == 'Elm Ave' ~ 33.86428452,
                               Monitor == 'North HS' ~ 33.83779526,
                               Monitor == 'G Street' ~ 33.77785794,
                               Monitor == 'Guenser Park' ~ 33.869231,
                               Monitor == 'Harbor Park' ~ 33.786252,
                               .default = latitude))
```



# Compute Daily Max H2S
```{r}
# Compute daily max
# -Inf (all NA) is converted to NA for that day
H2S_dm <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))

full_data <- full_data %>%
  left_join(H2S_dm, by = c('day', 'Monitor'))
```

```{r}
H2S_ma <- H2S_dm %>%
  mutate(yearmonth = format(day, "%Y-%m")) %>%
  group_by(yearmonth, Monitor) %>%
  summarise(H2S_monthly_average = mean(H2S_daily_max, na.rm = TRUE))

full_data <- full_data %>%
  left_join(H2S_ma, by = c('yearmonth', 'Monitor'))
```

## Compute daily average H2S
```{r}
# Compute daily average
H2S_da <- full_data %>%
  group_by(day, Monitor) %>%
  summarise(H2S_daily_avg = mean(H2S, na.rm=TRUE))

full_data <- full_data %>%
  left_join(H2S_da, by = c('day', 'Monitor'))
```

## Compute wind sector
```{r}
full_data <- full_data %>%
  mutate(wd_sec = case_when(0 <= wd & wd < 90 ~ 'Q1',
                            90 <= wd & wd < 180 ~ 'Q2',
                            180 <= wd & wd < 270 ~ 'Q3',
                            270 <= wd & wd <= 360 ~ 'Q4'))
```

## Read shape file of water treatment plants
```{r}
la_wrp <- readOGR('shapefiles/LA_WRP.shp', verbose=FALSE)

CRS_UTM <- CRS("+proj=utm +zone=11 ellps=WGS84")
utm_coord <- spTransform(la_wrp, CRS_UTM)

la_wrp <- tibble(name = la_wrp$CWP_NAM, capacity = la_wrp$cpcty_m, 
                 wrp_merc_x = coordinates(la_wrp)[,c(1)], 
                 wrp_merc_y = coordinates(la_wrp)[,c(2)],
                 wrp_utm_x =  coordinates(utm_coord)[,c(1)],
                 wrp_utm_y =  coordinates(utm_coord)[,c(2)])

monitors <- full_data %>%
  select(Monitor, latitude, longitude) %>%
  distinct()

monitors <- cbind(monitors, 
                  as.data.frame(project(as.matrix(cbind(monitors$longitude,
                                                        monitors$latitude)), 
                                        proj= "+proj=utm +zone=11 ellps=WGS84"))) %>%
  rename(mon_utm_x = V1,
         mon_utm_y = V2)

monitor_wrp <- monitors %>%
  full_join(la_wrp, by=character(), suffix = c(".m", ".w"))

monitor_wrp <- monitor_wrp %>%
  mutate(dist_wrp = sqrt((mon_utm_x - wrp_utm_x)^2 + (mon_utm_y - wrp_utm_y)^2))

monitor_wrp <- monitor_wrp %>%
  select(Monitor, dist_wrp, everything()) %>%
  arrange(Monitor) %>%
  group_by(Monitor) %>%
  slice_min(order_by = dist_wrp)

full_data <- full_data %>%
  left_join(monitor_wrp, join_by('Monitor'))
```

## Well production data
```{r}
well_prod <- read_csv('data/Los Angeles Production Well Monthly Production.CSV', show_col_types = FALSE)
well_info <- read_csv('data/Los Angeles Wells Well Headers.CSV', show_col_types = FALSE)
```

```{r}
well_prod <- well_prod %>%
  left_join(well_info, join_by(`API/UWI` == API14))

well_prod <- well_prod %>%
  select(`API/UWI`, `Monthly Production Date`, `Monthly Oil`, `Monthly Gas`, 
         `Monthly Water`, `Monthly BOE`, Days, `Operator Company Name.x`, 
         `Production Type.x`, `Well Status.x`, `Drill Type`, 
         `Surface Hole Latitude (WGS84)`, `Surface Hole Longitude (WGS84)`) %>%
  rename(lon = `Surface Hole Longitude (WGS84)`,
         lat = `Surface Hole Latitude (WGS84)`)

# add the projected sampler locations
well_location <- cbind(well_prod, 
                  as.data.frame(project(as.matrix(cbind(well_prod$lon,
                                                        well_prod$lat)), 
                                        proj= "+proj=utm +zone=11 ellps=WGS84"))) %>%
  select(`API/UWI`, lon, lat, V1, V2) %>%
  rename(well_lon = lon,
         well_lat = lat,
         well_utm_x = V1,
         well_utm_y = V2) %>%
  distinct()

monitor_well <- monitors %>%
  full_join(well_location, by=character(), suffix = c(".m", ".w"))

monitor_well <- monitor_well %>%
  mutate(distance = sqrt((mon_utm_x - well_utm_x)^2 + (mon_utm_y - well_utm_y)^2),
         `1km` = ifelse(distance<=1000, TRUE, FALSE),
         `2p5km` = ifelse(distance<=2500, TRUE, FALSE),
         `5km` = ifelse(distance<=5000, TRUE, FALSE))

# Keep closest
monitor_well <- monitor_well %>%
  select(Monitor, `API/UWI`, distance, everything()) %>%
  arrange(`API/UWI`) %>%
  group_by(`API/UWI`) %>%
  slice_min(order_by = distance)

well_prod <- well_prod %>%
  left_join(monitor_well, join_by(`API/UWI`))
```

```{r}
prod_data <- list(
    well_prod %>% 
      filter(`1km` == TRUE) %>% 
      group_by(`Monthly Production Date`) %>% 
      summarise(active_1km = n(),
                monthly_oil_1km = sum(`Monthly Oil`, na.rm = TRUE),
                monthly_gas_1km = sum(`Monthly Gas`, na.rm = TRUE),
                monthly_water_1km = sum(`Monthly Water`, na.rm = TRUE),
                monthly_boe_1km = sum(`Monthly BOE`, na.rm = TRUE))
    ,
    well_prod %>% 
      filter(`2p5km` == TRUE) %>% 
      group_by(`Monthly Production Date`) %>% 
      summarise(active_2p5km = n(),
            monthly_oil_2p5km = sum(`Monthly Oil`, na.rm = TRUE),
            monthly_gas_2p5km = sum(`Monthly Gas`, na.rm = TRUE),
            monthly_water_2p5km = sum(`Monthly Water`, na.rm = TRUE),
            monthly_boe_2p5km = sum(`Monthly BOE`, na.rm = TRUE))
    ,
    well_prod %>% 
      filter(`5km` == TRUE) %>% 
      group_by(`Monthly Production Date`) %>% 
      summarise(active_5km = n(),
            monthly_oil_5km = sum(`Monthly Oil`, na.rm = TRUE),
            monthly_gas_5km = sum(`Monthly Gas`, na.rm = TRUE),
            monthly_water_5km = sum(`Monthly Water`, na.rm = TRUE),
            monthly_boe_5km = sum(`Monthly BOE`, na.rm = TRUE))
  ) %>%
  reduce(inner_join, by = 'Monthly Production Date') %>%
  mutate(yearmonth = format(`Monthly Production Date`, '%Y-%m'))

full_data <- full_data %>%
  left_join(prod_data, join_by(yearmonth))
```


```{r}
saveRDS(full_data, 'data/full_data.rds')
```

# Other issues
- Seems like for Torrance Refinery, the three monitors had different locations recorded at first, but stayed the same for many observations.
