---
title: "Merge LA refinery data sets"
author: "Jerry Wu"
date: "9/01/2024"
output:
  html_document:
    df_print: paged
abstract: ''
subtitle: ''
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(plotly)
library(units)
library(lwgeom)
library(circular)
library(raster)
library(units)
library(lubridate)
select <- dplyr::select
```

```{r}
# load data sets
full_data <- tibble('Date/Time' = character(),
                    'H2S' = numeric(),
                    'Wind Direction' = numeric(),
                    'Wind Speed' = numeric(),
                    'Ammonia' = numeric(),
                    'Benzene' = numeric(),
                    'Black Carbon' = numeric())

for (file in list.files('../data/new_data_all')){
  site <- sub('_All.csv', '', file)
  new_data <- read_csv(paste0('../data/new_data_all/', file))
  new_data$Monitor <- site
  print(names(new_data))
  full_data <- bind_rows(full_data, new_data)
}

remove(new_data)
gc()
```

```{r}
# check duplicate date/site pairs
# full_data %>% 
#   select(`Date/Time`, Monitor, DST) %>%
#   group_by(`Date/Time`, Monitor, DST) %>%
#   summarise(n = n()) %>%
#   filter(n > 2) %>%
#   nrow()
```

```{r}
site_geo_metadata <- read_csv('../data/site_geo_metadata.csv')

full_data <- full_data %>%
  left_join(site_geo_metadata, join_by('Monitor' == 'site_name'))
```


# Compute date variables
```{r}
full_data <- full_data %>% 
  mutate(`Date/Time` = as.POSIXct(`Date/Time`, "%m/%d/%Y %H:%M", tz = 'America/Los_Angeles'))

full_data <- full_data %>% 
  mutate(day = as.POSIXct(format(`Date/Time`, "%Y-%m-%d"), format="%Y-%m-%d", tz = 'America/Los_Angeles'))
 
# Remove observations before 2020
full_data <- full_data %>% filter(day > '2019-12-31') %>% distinct()
 
full_data <- full_data %>% 
  mutate(yearmonth = format(day, "%Y-%m"),
         year = format(day, "%Y"),
         month = format(day, "%m"),
         dayhour = format(`Date/Time`, "%Y-%m-%d %H"),
         weekday = relevel(factor(wday(full_data$day, label=TRUE), ordered = FALSE), ref = "Sun"))

# save for weather data download
# saveRDS(full_data, '../data/weather/site_data_for_weather.rds')
```

```{r}
# add one hour to date/time where DST is true
full_data <- full_data %>%
  mutate(`Date/Time` = if_else(DST == TRUE, `Date/Time` + hours(1), `Date/Time`))
```

```{r}
# check for duplicate time for each monitor
# full_data %>% 
#   select(`Date/Time`, Monitor) %>%
#   group_by(`Date/Time`, Monitor) %>%
#   summarise(n = n()) %>%
#   filter(n > 2) %>%
#   nrow()
```


# Add meteorological data
```{r}
# load in daily weather data
monitor_daily_weather <- readRDS('../data/weather/monitor_daily_weather.rds')
monitor_hourly_weather <- readRDS('../data/weather/monitor_hourly_weather.rds')
```

```{r}
# Merge with full_data using day and Monitor as keys for daily data
# prepare daily weather
monitor_daily_weather <- monitor_daily_weather %>% 
  mutate(day = as.POSIXct(day, format="%Y-%m-%d", tz = 'America/Los_Angeles')) %>% 
  select(day, Monitor, daily_temp, daily_hum, daily_precip, daily_ws, daily_wd)

full_data <- full_data %>%
  left_join(monitor_daily_weather, join_by(Monitor, day))
```

```{r}
# same for hourly data
# first prepare hourly weather data
monitor_hourly_weather <- monitor_hourly_weather %>% 
  mutate(`Date/Time` = as.POSIXct(hourly, format = "%Y-%m-%dT%H:%M", tz = 'America/Los_Angeles')) %>%
  mutate(dayhour = format(`Date/Time`, "%Y-%m-%d %H"))  %>% 
  select(dayhour, Monitor, hourly_temp, hourly_hum, hourly_precip, hourly_ws, hourly_wd)

full_data <- full_data %>%
  left_join(monitor_hourly_weather, join_by(Monitor, dayhour), keep = FALSE)
```

```{r}
# We also want to fill missing wind data for the monitors without wind measurements
# !!! As for now, this will be done using daily wind data instead of hourly due to DST issues
full_data <- full_data %>%
  mutate(`Wind Speed` = coalesce(`Wind Speed`, daily_ws),
         `Wind Direction` = coalesce(`Wind Direction`, daily_wd))
```


<!-- # This part was used to add weather data -->
<!-- # Add wind data for El Segundo and St Anthony -->
<!-- ```{r} -->
<!-- klax_data <- read_csv('data/KLAX_2020-2023.csv', show_col_types = FALSE) %>%  -->
<!--   filter(!row_number() == 1) -->

<!-- klax_daily <- klax_data %>% -->
<!--   group_by(Date) %>% -->
<!--   summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T), -->
<!--             avg_hum = mean(as.numeric(relative_humidity), na.rm=T), -->
<!--             precip = precip_accum_24_hour[which(!is.na(precip_accum_24_hour))[1]], -->
<!--             ws = mean(as.numeric(wind_speed), na.rm=TRUE), -->
<!--             wd = as.numeric(mean(circular(as.numeric(wind_direction), units = 'degrees'), na.rm=TRUE))) %>% -->
<!--   mutate(precip = coalesce(as.numeric(precip), 0), -->
<!--          Date = parse_date_time(Date, c('m/d/y')), -->
<!--          Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC'), -->
<!--          wd = if_else(wd < 0, wd + 360, wd)) -->

<!-- full_data <- rbind(full_data %>%  -->
<!--                      filter(!(Monitor %in% c('El Segundo', 'St. Anthony'))),  -->
<!--                    full_data %>% -->
<!--                      select(-c(wd, ws)) %>% -->
<!--                      filter(Monitor %in% c('El Segundo', 'St. Anthony')) %>% -->
<!--                      left_join(klax_daily %>% select(Date, ws, wd), join_by(day == Date)))  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # Remove rows without wd or ws -->
<!-- full_data <- full_data %>%  -->
<!--   drop_na(c("wd", "ws")) -->
<!-- ``` -->
<!-- # Weather data from National Weather Service -->
<!-- ```{r} -->
<!-- # klax_data <- read_csv('data/KLAX_2020-2023.csv', show_col_types = FALSE) %>%  -->
<!-- #   filter(!row_number() == 1) -->
<!-- ktoa_data <- read_csv('data/KTOA_2020-2023.csv', show_col_types = FALSE) %>%  -->
<!--   filter(!row_number() == 1) -->
<!-- klgb_data <- read_csv('data/KLGB_2020-2023.csv', show_col_types = FALSE) %>%  -->
<!--   filter(!row_number() == 1) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # compute daily average temperature, humidity, and precipitation -->
<!-- # klax_daily <- klax_data %>% -->
<!-- #   group_by(Date) %>% -->
<!-- #   summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T), -->
<!-- #             avg_hum = mean(as.numeric(relative_humidity), na.rm=T), -->
<!-- #             precip = precip_accum_24_hour[which(!is.na(precip_accum_24_hour))[1]]) %>% -->
<!-- #   mutate(precip = coalesce(as.numeric(precip), 0), -->
<!-- #          Date = parse_date_time(Date, c('m/d/y')), -->
<!-- #          Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC')) -->

<!-- klax_daily <- klax_daily %>%  -->
<!--   select(-c(ws, wd)) -->

<!-- klgb_daily <- klgb_data %>% -->
<!--   group_by(Date) %>% -->
<!--   summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T), -->
<!--             avg_hum = mean(as.numeric(relative_humidity), na.rm=T), -->
<!--             precip = precip_accum_24_hour[which(!is.na(precip_accum_24_hour))[1]]) %>% -->
<!--   mutate(precip = coalesce(as.numeric(precip), 0), -->
<!--          Date = parse_date_time(Date, c('m/d/y')), -->
<!--          Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC')) -->

<!-- ktoa_daily <- ktoa_data %>% -->
<!--   group_by(Date) %>% -->
<!--   summarise(avg_temp = mean(as.numeric(air_temp), na.rm=T), -->
<!--             avg_hum = mean(as.numeric(relative_humidity), na.rm=T)) %>% -->
<!--   mutate(Date = parse_date_time(Date, c('m/d/y')), -->
<!--          Date = as.POSIXct(Date, "%Y-%m-%d", tz = 'UTC')) -->

<!-- remove(klax_data, ktoa_data, klgb_data) -->

<!-- # notice that the Torrance station is missing 10 days of observations -->
<!-- # for that 10 days, use the long beach airport data as substitutes -->
<!-- ktoa_daily <- bind_rows(ktoa_daily,  -->
<!--                         anti_join(klgb_daily,ktoa_daily, by = 'Date') %>% select(-precip)) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # map each monitor to closest monitor -->
<!-- weather_stations <- tibble(station = c('KLAX', 'KLGB', 'KTOA'), -->
<!--                            latitude = c(33.93806, 33.81167, 33.803979), -->
<!--                            longitude = c(-118.33194, -118.38889, -118.339432)) %>% -->
<!--   st_as_sf(coords = c('longitude', 'latitude')) %>%  -->
<!--   st_set_crs(4326) -->

<!-- weather_stations <- st_transform(weather_stations, CRS_UTM) -->

<!-- closest_temp_hum_station <- weather_stations$station[st_nearest_feature(monitors, weather_stations)] -->

<!-- closest_precip_station <- weather_stations$station[st_nearest_feature(monitors, weather_stations %>% filter(station != 'KTOA'))] -->

<!-- monitors_weather <- tibble(Monitor = monitors$Monitor,  -->
<!--                            weather_station = closest_temp_hum_station, -->
<!--                            closest_precip_station = closest_precip_station) -->

<!-- weather_full <- bind_rows(klax_daily %>% select(-precip) %>%  -->
<!--                             mutate(weather_station = 'KLAX'), -->
<!--                           klgb_daily %>% select(-precip) %>%  -->
<!--                             mutate(weather_station = 'KLGB'), -->
<!--                           ktoa_daily %>%  -->
<!--                             mutate(weather_station = 'KTOA')) -->

<!-- precip_full <- bind_rows(klax_daily %>% select(Date, precip) %>%  -->
<!--                             mutate(closest_precip_station = 'KLAX'), -->
<!--                           klgb_daily %>% select(Date, precip) %>%  -->
<!--                             mutate(closest_precip_station = 'KLGB')) -->

<!-- full_data <- full_data %>% -->
<!--   left_join(monitors_weather, join_by('Monitor')) %>% -->
<!--   left_join(weather_full, join_by('day' == 'Date', 'weather_station')) %>% -->
<!--   left_join(precip_full, join_by('day' == 'Date', 'closest_precip_station')) -->

<!-- remove(klax_daily, ktoa_daily, klgb_daily, precip_full, weather_full,  -->
<!--        monitors_weather, weather_stations) -->
<!-- gc() -->
<!-- ``` -->


```{r}
# Check for duplicate time
# full_data %>%
#    group_by(DateTime, Monitor) %>%
#    summarise(n = n()) %>%
#    filter(n > 1)
```
We observe multiple rows for the same date time and monitor. 
From a quick glance, this could be a conversion issue from Date.Time to DateTime, since the H2S measurements are different. 
In other occasions, multiple rows exist as complements, where one row contains data on the missing values of the other.
It also appears that some times we get different measurements for the same time when it's not daylight saving conversion days?

- example: January 31, G Street, wd and ws are recorded more than once.
```{r}
# full_data %>%
#     group_by(DateTime, Monitor) %>%
#     summarise(n = n()) %>%
#     filter(n > 1) %>%
#     mutate(day = as.POSIXct(format(DateTime, "%Y-%m-%d"), format="%Y-%m-%d", tz = 'UTC')) %>%
#     group_by(day, Monitor) %>%
#     summarise(n = n())
```


```{r}
# Deal with duplicate rows for same time

# For the 213 monitor it seems like the duplicate rows are complements to the other
#https://stackoverflow.com/questions/45515218/combine-rows-in-data-frame-containing-na-to-make-complete-row
# coalesce_by_column <- function(df) {
#   return(dplyr::coalesce(!!! as.list(df)))
# }
# 
# data_213 <- full_data %>%
#   filter(Monitor == '213th & Chico') %>%
#   group_by(DateTime) %>%
#   summarise_all(coalesce_by_column)
# 
# full_data <- full_data %>% 
#   filter(Monitor != '213th & Chico') %>%
#   rbind(data_213)
```


```{r}
#glimpse(full_data)
```

```{r}
#summary(full_data)
```
Some concerns for the data include:
For TorranceAir_H2SWind, there are actually three sites, but only one site contains ws/wd information
What's the difference between DateTime and Date/Time?
DateTime would be the time after converting
What's the difference between Date/Time and Date/Time.First? The latter is present in FirstMethodist
For missing H2S, it could be missing due to mismatch from merging, or flagged due to issues such as low detection (minimm detection limit: 0.4)

# Any additional cleaning
```{r}
# For observations below MDL, set it to half of MDL
# As confirmed, MDL for all monitors is 0.4
# Check how many observations are below MDL
full_data %>%
  group_by() %>%
  summarise('Below MDL' = sum(full_data$H2S < 0.4, na.rm=T),
            'Proportion' = sum(full_data$H2S < 0.4, na.rm=T) / n())

full_data <- full_data %>%
  mutate(H2S = if_else(H2S < 0.4, 0.2, H2S))
```

# Compute Hourly average H2S
```{r}
# Compute hourly average
# -Inf (all NA) is converted to NA for that day
H2S_ha <- full_data %>%
  group_by(dayhour, Monitor, DST) %>%
  summarise(H2S_hourly_avg = mean(H2S, na.rm = TRUE)) %>%
  mutate(H2S_hourly_avg = if_else(H2S_hourly_avg == -Inf, NA, H2S_hourly_avg))

full_data <- full_data %>%
  left_join(H2S_ha, join_by(dayhour, Monitor, DST))
```

# Compute Hourly max H2S
```{r}
# Compute hourly max
# -Inf (all NA) is converted to NA for that day
H2S_hm <- full_data %>%
  group_by(dayhour, Monitor, DST) %>%
  summarise(H2S_hourly_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_hourly_max = if_else(H2S_hourly_max == -Inf, NA, H2S_hourly_max))

full_data <- full_data %>%
  left_join(H2S_hm, join_by(dayhour, Monitor, DST))
```

# Compute daily average H2S
```{r}
# Compute daily average
H2S_da <- full_data %>%
  group_by(day, Monitor, DST) %>%
  summarise(H2S_daily_avg = mean(H2S, na.rm=TRUE))

full_data <- full_data %>%
  left_join(H2S_da, by = join_by(day, Monitor, DST))
```

# Compute Daily Max H2S
```{r}
# Compute daily max
# -Inf (all NA) is converted to NA for that day
H2S_dm <- full_data %>%
  group_by(day, Monitor, DST) %>%
  summarise(H2S_daily_max = max(H2S, na.rm = TRUE)) %>%
  mutate(H2S_daily_max = if_else(H2S_daily_max == -Inf, NA, H2S_daily_max))

full_data <- full_data %>%
  left_join(H2S_dm, by = join_by(day, Monitor, DST))
```
```{r}
remove('H2S_ha', 'H2S_hm', 'H2S_da', 'H2S_dm')
```

<!-- # Compute Monthly Average H2S -->
<!-- ```{r} -->
<!-- H2S_ma <- H2S_dm %>% -->
<!--   mutate(yearmonth = format(day, "%Y-%m")) %>% -->
<!--   group_by(yearmonth, Monitor) %>% -->
<!--   summarise(H2S_monthly_average = mean(H2S_daily_max, na.rm = TRUE)) -->

<!-- full_data <- full_data %>% -->
<!--   left_join(H2S_ma, by = c('yearmonth', 'Monitor')) -->

<!-- remove(H2S_ma, H2S_dm) -->
<!-- ``` -->



# Compute wind sector
```{r}
full_data <- full_data %>%
  mutate(wd_sec = case_when(0 <= `Wind Direction` & `Wind Direction` < 90 ~ 'Q1',
                            90 <= `Wind Direction` & `Wind Direction` < 180 ~ 'Q2',
                            180 <= `Wind Direction` & `Wind Direction` < 270 ~ 'Q3',
                            270 <= `Wind Direction` & `Wind Direction` <= 360 ~ 'Q4'))
```

# Well production data
```{r}
well_prod <- read_csv('../data/LA well prod Jun2024 Well Monthly Production.CSV', show_col_types = FALSE)
full_well_info <- read_csv('../data/well_active_inactive_prod_since_2000.CSV', show_col_types = FALSE)
active_well_info <- read_csv('../data/LA well prod Jun2024 Well Headers.CSV', show_col_types = FALSE)
```

```{r}
CRS_UTM <- CRS("+proj=utm +zone=11 ellps=WGS84")

monitors <- full_data %>%
  select(Monitor, lat, lon) %>%
  distinct() %>%
  st_as_sf(coords = c('lon', 'lat')) %>%
  st_set_crs(4326) %>%
  st_transform(CRS_UTM)

inactive_well_info <- anti_join(full_well_info, active_well_info, join_by(API14)) %>%
  rename(lon = `Surface Hole Longitude (WGS84)`,
         lat = `Surface Hole Latitude (WGS84)`) %>%
  select(`API14`, lon, lat) %>%
  distinct() %>%
  st_as_sf(coords = c('lon', 'lat')) %>% 
  st_set_crs(4326) %>%
  st_transform(CRS_UTM) 

well_prod <- well_prod %>%
  left_join(active_well_info, join_by(`API/UWI` == API14)) 

well_prod <- well_prod %>%
  select(`API/UWI`, `Monthly Production Date`, `Monthly Oil`, `Monthly Gas`, 
         `Monthly Water`, `Monthly BOE`, Days, `Operator Company Name.x`, 
         `Production Type.x`, `Well Status.x`, `Drill Type`, 
         `Surface Hole Latitude (WGS84)`, `Surface Hole Longitude (WGS84)`) %>%
  rename(lon = `Surface Hole Longitude (WGS84)`,
         lat = `Surface Hole Latitude (WGS84)`)

# project distinct well locations to UTM
active_well_location <- well_prod %>%
  select(`API/UWI`, lon, lat) %>%
  distinct() %>%
  st_as_sf(coords = c('lon', 'lat')) %>% 
  st_set_crs(4326) %>%
  st_transform(CRS_UTM) 

# distance to closest monitor for each well
monitor_active_wells <- tibble()
for (i in 1:nrow(monitors)){
  monitor_well_api <- drop_units(st_distance(monitors[i, ], active_well_location))
  monitor_well_api <- active_well_location[which(monitor_well_api <= 2000), ] %>% st_drop_geometry()
  monitor_active_wells <- rbind(monitor_active_wells, 
                                tibble(Monitor = monitors[i, 1]$Monitor, 
                                       API14 = monitor_well_api$`API/UWI`))
}

monitor_well_prod <- monitor_active_wells %>% 
  left_join(well_prod, join_by(API14 == `API/UWI`)) %>%
  group_by(Monitor, `Monthly Production Date`) %>%
  summarise(active_2km = n(),
            monthly_oil_2km = sum(`Monthly Oil`, na.rm = TRUE),
            monthly_gas_2km = sum(`Monthly Gas`, na.rm = TRUE)) %>%
  mutate(yearmonth = format(`Monthly Production Date`, '%Y-%m')) %>%
  select(-`Monthly Production Date`)

full_data <- full_data %>%
  left_join(monitor_well_prod, join_by(Monitor, yearmonth)) %>%
  mutate(active_2km = coalesce(active_2km, 0),
            monthly_oil_2km = coalesce(monthly_oil_2km, 0),
            monthly_gas_2km = coalesce(monthly_gas_2km, 0))

monitor_inactive_wells <- tibble()
for (i in 1:nrow(monitors)){
  monitor_well_api <- drop_units(st_distance(monitors[i, ], inactive_well_info))
  monitor_well_api <- inactive_well_info[which(monitor_well_api <= 2000), ] %>% st_drop_geometry()
  monitor_inactive_wells <- rbind(monitor_inactive_wells, 
                                  tibble(Monitor = monitors[i, 1]$Monitor, 
                                         inactive_2km = nrow(monitor_well_api)))
}

full_data <- full_data %>%
  left_join(monitor_inactive_wells, join_by(Monitor))
```

# Downwind Indicator wrp    
```{r}
# Create variable to indicate downwind wrt refinery
wind_diff <- abs(full_data$angle_ref - 180 - full_data$`Wind Direction`)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
full_data$downwind_ref <- as.integer(wind_diff <= 15)
```

# Downwind of wrp
```{r}
# Finally, check if wind direction is downwind w.r.t wrp
wind_diff <- abs(full_data$angle_wrp - full_data$`Wind Direction`)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

full_data$downwind_wrp <- as.integer(wind_diff <= 15)
```

# Compute UTM coordinates for monitors
```{r}
CRS_UTM <- CRS("+proj=utm +zone=11 ellps=WGS84")

monitors <- monitors %>%
  mutate(mon_utm_x = st_coordinates(.)[,1],
         mon_utm_y = st_coordinates(.)[,2]) %>%
  st_drop_geometry
                   

full_data <- full_data %>%
  left_join(monitors, join_by(Monitor))
```

```{r}
# Compute daily data
daily_full <- full_data %>%
  select(H2S_daily_max, H2S_daily_avg, `Wind Speed`, `Wind Direction`,
         lat, lon, mon_utm_x, mon_utm_y, Monitor, dist_ref, 
         angle_ref, closest_ref, year, month, day, weekday, 
         active_2km, monthly_oil_2km, monthly_gas_2km, inactive_2km,
         dist_wrp, closest_wrp_capacity, angle_wrp, dist_dc,
         daily_temp, daily_hum, daily_precip) %>%
  group_by(Monitor, day) %>%
  mutate(ws_avg = mean(`Wind Speed`, na.rm=TRUE),
         wd_avg = as.numeric(mean(circular(`Wind Direction`, units = 'degrees'), na.rm=TRUE))) %>%
  mutate(wd_avg = if_else(wd_avg < 0, wd_avg+360, wd_avg)) %>%
  ungroup() %>%
  rename(monitor_lat = lat, monitor_lon = lon)  %>%
  select(-c(`Wind Speed`, `Wind Direction`)) %>%
  distinct()
```

```{r}
# Get the downwind indicators for daily data
wind_diff <- abs(daily_full$angle_ref - 180 - daily_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
daily_full$daily_downwind_ref <- as.integer(wind_diff <= 15)

wind_diff <- abs(daily_full$angle_wrp - daily_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

daily_full$daily_downwind_wrp <- as.integer(wind_diff <= 15)
```

```{r}
# Compute hourly data
hourly_full <- full_data %>%
  select(H2S_hourly_max, H2S_hourly_avg, `Wind Speed`, `Wind Direction`,
         lat, lon, mon_utm_x, mon_utm_y, Monitor, dist_ref,
         angle_ref, closest_ref, year, month, day, weekday, dayhour,
         active_2km, monthly_oil_2km, monthly_gas_2km, inactive_2km,
         dist_wrp, closest_wrp_capacity, angle_wrp, dist_dc,
         hourly_temp, hourly_precip, hourly_hum) %>%
  group_by(Monitor, dayhour) %>%
  mutate(ws_avg = mean(`Wind Speed`, na.rm=TRUE),
         wd_avg = as.numeric(mean(circular(`Wind Direction`, units = 'degrees'), na.rm=TRUE))) %>%
  mutate(wd_avg = if_else(wd_avg < 0, wd_avg+360, wd_avg)) %>%
  ungroup() %>%
  rename(monitor_lat = lat, monitor_lon = lon)  %>%
  select(-c(`Wind Speed`, `Wind Direction`)) %>%
  distinct()
```

```{r}
# Get the downwind indicators for hourly data
wind_diff <- abs(hourly_full$angle_ref - 180 - hourly_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)
hourly_full$hourly_downwind_ref <- as.integer(wind_diff <= 15)

wind_diff <- abs(hourly_full$angle_wrp - hourly_full$wd_avg)
wind_diff <- ifelse(wind_diff > 180, 360 - wind_diff, wind_diff)

hourly_full$hourly_downwind_wrp <- as.integer(wind_diff <= 15)
```

# Get monitor locations
```{r}
monitors_coord <- full_data %>%
  select(Monitor, lat, lon) %>%
  distinct()

monitors <- monitors_coord %>%
  select(Monitor)
  
coordinates(monitors_coord) <- ~ lon + lat
```

# Load in elevation data
```{r}
elevation <- raster('../shapefiles/N33W119.hgt')

monitors <- cbind(monitors, extract(elevation, monitors_coord)) %>%
  as.data.frame() %>%
  rename(elevation = 2)
```

# Load in EVI data
```{r}
#evi <- raster('shapefiles/MOD13Q1.061__250m_16_days_EVI_doy2023145_aid0001.tif')
evi <- raster('../shapefiles/MOD13Q1.006__250m_16_days_EVI_doy2022177_aid0001.tif')
monitors <- cbind(monitors, extract(evi, monitors_coord) * 0.0001) %>%
  as.data.frame() %>%
  rename(EVI = 3)
```

# Merge EVI and elevation to data
```{r}
full_data <- full_data %>%
  left_join(monitors, join_by(Monitor))

hourly_full <- hourly_full %>%
  left_join(monitors, join_by(Monitor))

daily_full <- daily_full %>%
  left_join(monitors, join_by(Monitor))
```

# Merge Odor Complaint data
```{r}
odor <- read_csv('../data/odorcomplaintdata_2018_2023.csv')
la_county <- read_sf('../shapefiles/Zip_Codes_(LA_County)/Zip_Codes_(LA_County).shp')
```

```{r}
# https://gis.stackexchange.com/questions/282750/identify-polygon-containing-point-with-r-sf-package
pnts_sf <- do.call("st_sfc",c(lapply(1:nrow(monitors_coord@coords), 
function(i) {st_point(as.numeric(monitors_coord@coords[i, ]))}), list("crs" = 4326))) 

pnts_trans <- st_transform(pnts_sf, CRS_UTM) # apply transformation to pnts sf
la_county_trans <- st_transform(la_county, CRS_UTM)      # apply transformation to polygons sf

# intersect and extract state name
monitors_coord@data$county <- apply(st_intersects(la_county_trans, pnts_trans, sparse = FALSE), 2, 
               function(col) { 
                  la_county_trans[which(col), ]$ZIPCODE
               })
```

```{r}
library(ggrepel)
# Plot results to double check
la_county_present <- la_county[la_county$ZIPCODE %in% unique(monitors_coord$county), ]

monitor_zip <- tibble(Monitor = monitors_coord@data$Monitor,
                      zip = monitors_coord@data$county,
                      longitude = monitors_coord@coords[,1],
                      latitude = monitors_coord@coords[,2])

monitor_zip_graph <- ggplot() +
  geom_sf(data = la_county_present, aes(fill = ZIPCODE)) +
  geom_point(data = monitor_zip, 
             aes(x = longitude, y = latitude))

monitor_zip_graph + 
  geom_label_repel(data = monitor_zip, aes(x = longitude, y = latitude, label = Monitor),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50') +
  theme_minimal()
```

```{r}
odor <- odor[odor$zip %in% unique(monitors_coord$county), ]

monitor_odor <- monitors_coord@data %>%
  left_join(odor %>% mutate(zip = as.character(zip)), join_by(county == zip)) %>%
  select(-`...1`)

full_data <- full_data %>%
  left_join(monitors_coord@data, join_by(Monitor))

hourly_full <- hourly_full %>%
  left_join(monitors_coord@data, join_by(Monitor))

daily_full <- daily_full %>%
  left_join(monitors_coord@data, join_by(Monitor))

full_data <- full_data %>%
  left_join(monitor_odor %>% select(Monitor, date, num_odor_complaints), join_by(Monitor, day == date))

full_data <- full_data %>%
  mutate(num_odor_complaints = coalesce(num_odor_complaints, 0))

hourly_full <- hourly_full %>%
  left_join(monitor_odor %>% select(Monitor, date, num_odor_complaints), join_by(Monitor, day == date))

hourly_full <- hourly_full %>%
  mutate(num_odor_complaints = coalesce(num_odor_complaints, 0))

daily_full <- daily_full %>%
  left_join(monitor_odor %>% select(Monitor, date, num_odor_complaints), join_by(Monitor, day == date))

daily_full <- daily_full %>%
  mutate(num_odor_complaints = coalesce(num_odor_complaints, 0))
```

```{r}
odor_graph <- odor %>%
  filter(zip %in% unique(monitors_coord$county)) %>%
  mutate(zip = as.character(zip)) %>%
   ggplot(aes(x=date, y=num_odor_complaints, group=zip, color=zip)) +
   geom_line() +
   labs(title = "Number of odor complaints over time", y = 'Number of odor complaints', x = 'time') +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(odor_graph) %>% layout(dragmode = 'pan')
```

```{r}
odor_graph_b <- odor %>%
  filter(zip %in% unique(monitors_coord$county)) %>%
  mutate(zip = as.character(zip)) %>%
   ggplot(aes(x=date, y=num_odor_complaints, group=zip, color=zip)) +
   geom_line() +
   labs(title = "Number of odor complaints over time w/o outliers", y = 'Number of odor complaints', x = 'time') +
   scale_y_continuous(limits = c(0, 30)) +  
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45))

ggplotly(odor_graph_b) %>% layout(dragmode = 'pan')
```

```{r}
# save data until September 2023 due to missing odor data
full_data <- full_data %>%
  filter(day < '2023-10-01')

hourly_full <- hourly_full %>%
  filter(day < '2023-10-01')

daily_full <- daily_full %>%
  filter(day < '2023-10-01')
```

```{r}
saveRDS(full_data, '../data/full_data_20230930.rds')
saveRDS(hourly_full, '../data/hourly_full_20230930.rds')
saveRDS(daily_full, '../data/daily_full_20230930.rds')
```

```{r}
remove(la_county, la_county_present, la_county_trans, monitor_odor, monitor_zip,
       monitor_zip_graph, monitors, monitors_coord, odor, odor_graph, odor_graph_b,
       pnts_sf, pnts_trans, elevation, evi, active_well_info, active_well_location,
       full_well_info, inactive_well_info, monitor_daily_weather, monitor_hourly_weather,
       monitor_inactive_wells, monitor_well_api, monitor_well_prod, site_geo_metadata,
       well_prod, monitor_active_wells)
gc()
```


# Other issues
- Seems like for Torrance Refinery, the three monitors had different locations recorded at first, but stayed the same for many observations.
